### 2024-10-22 paper/paper.qmd
Initial Submission
First Draft

---
title: "Forcasting US Election 2024"
subtitle: "##TODO: result here"
author: 
  - Maryam Ansari
  - Amy Jin
  - Maggie Zhang

thanks: "Code and data are available at: [https://github.com/aj3616/Forcasting-US-Elections](https://github.com/aj3616/Forcasting-US-Elections)."
date: today
date-format: long
abstract: "first sentence: specify the general area of the paper and encourage the reader; second sentence: specify the dataset and methods at a general level; third sentence: specify the headline result; and a fourth sentence about implications"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(ggplot2)
library(tidyr)
data = read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
```

# Introduction

As of October-November 2024, the U.S. presidential election is almost at its final stage, with all candidates actively campaigning across numerous states. The presidential election will happen on November 5, 2024 (@calendar). The two major parties, Democrats and Republicans, are in particular focus. Former President Donald Trump represents the Republican party, and President Joe Biden initially led re-election for the Democrats. However, Biden has dropped out, and Kamala Harris has taken over as the Democratic nominee (@g24), who is the first Black woman nominee. Harris is the Both candidates are focusing their efforts on securing votes by addressing their aspect on social issues. pollsters and election forecasting with data evidence are important in predicting election outcomes. Polling can highlight underlying or emerging issues, as well as reveal voter preferences. The polling outcomes allow campaigns of different parties to strategically target specific demographics. For example in 2012, pollsters and data modellers such as Nate Silver have all used survey research and statistical models to successfully predict the result of Barack Obama's victory in the presidential election with consistent and reliable forecasts (@b17).

The remainder of this paper is structured as follows. @sec-data....

# Polling Data {#sec-data}

## Overview

We use the polling data to forecast the potential outcomes of the 2024 U.S. presidential election between Kamala Harris and Donald Trump.The dataset was obtained from https://projects.fivethirtyeight.com/polls/president-general/2024/national/ or (FiveThirtyEight’s “Poll of Polls” for the 2024 U.S. Presidential election [@fivethirtyeight]). It provides a comprehensive view of voter preferences through aggregated results from numerous national polls conducted by various polling organizations. The dataset was simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2].

## Variables

The dataset comprises several key variables of interest, including poll_id, a unique identifier for each poll conducted; pollster, which indicates the organization conducting the poll; sample_size, representing the total number of respondents; population, specifying the voting group described (e.g., likely voters); candidate_name, the names of the candidates in the poll (e.g., Kamala Harris, Donald Trump); and pct, the percentage of the vote or support received by each candidate. These variables allow us to explore various dimensions of polling data, such as trends in voter support across different states and the influence of pollster reliability on polling outcomes. The (@tbl-firstten) below provides a snapshot of the polling data, displaying the first ten entries. This includes the pollster names, sample sizes, and the percentage of support for each candidate, which can reveal patterns in public opinion and help identify how different organizations may report varying levels of support for Kamala Harris and Donald Trump.

```{r}
#| label: tbl-firstten
#| tbl-cap: "Sample sizes and support percentages for Kamala Harris and Donald Trump"
#| echo: false
# Drop rows with NA values and select only the relevant columns
firstten_data <- drop_na(data) %>%
  select(pollster_id, pollster, sample_size, population, candidate_name, pct)

# Get the first 10 entries
first_ten_entries <- firstten_data %>%
  slice(1:10)

# Create the gt table with only the selected columns
firstten <- first_ten_entries %>%
  gt() %>%
  tab_header(
    title = "Polling Data Snapshot: 2024 U.S. Presidential Election",
  ) %>%
  cols_label(
    pollster_id = "Poll ID",
    pollster = "Pollster",
    sample_size = "Sample Size",
    population = "Population",
    candidate_name = "Candidate",
    pct = "Support (%)"
  ) %>%
  fmt_number(
    columns = c(sample_size, pct),
    decimals = 0
  )

# Display the gt table
firstten

```

## Summary statistics & Relationships

### Pollster Reliability and Election Outcome

The relationship between polling organization reliability, as indicated by the pollscore, and the percentage of support (pct) for each candidate was examined. It is anticipated that more reliable pollsters will yield more accurate predictions. Consequently, polls were categorized into three tiers based on their pollscore: high reliability (pollscore greater than 0), medium reliability (pollscore between -1 and 0), and low reliability (pollscore less than -1). By comparing the average support percentages (pct) for Kamala Harris and Donald Trump across these tiers, this analysis sought to identify whether more reliable pollsters produce different outcomes than those deemed less reliable. Mean pct values were calculated for each category, and standard deviation was included to illustrate variability within the poll results. The table(@tbl-reliability) and bar chart(@fig-reliability) showing the average `pct` for each candidate across different levels of pollster reliability. Discussion: This analysis highlights whether pollsters with higher reliability scores offer more accurate predictions and if their estimates favor one candidate over the other.

```{r}
#| label: tbl-reliability
#| tbl-cap: "Mean and Standard Deviation of Support Percentages (pct) for Kamala Harris and Donald Trump Across Pollster Reliability Levels"
#| echo: false
# Categorizing poll reliability
# Create a new dataset with categorized reliability
# Remove rows with NA values in the 'pollscore' column while keeping the rest of the data frame
pollscore_data <- data %>%
  filter(!is.na(pollscore)) %>%
  mutate(
    reliability = case_when(
      pollscore > 0 ~ "High",
      pollscore > -1 & pollscore <= 0 ~ "Medium",
      pollscore <= -1 ~ "Low"
    )
  )

# Calculate mean and standard deviation for each candidate across reliability levels
summary_data <- pollscore_data %>%
  group_by(candidate_name, reliability) %>%
  summarise(
    mean_pct = mean(pct),
    sd_pct = sd(pct),
    .groups = 'drop'
  )
reliability_table <- summary_data %>%
  gt() %>%
  tab_header(
    title = "Polling Data Summary: Pollster Reliability and Support Percentages"
  ) %>%
  cols_label(
    candidate_name = "Candidate",
    reliability = "Pollster Reliability",
    mean_pct = "Mean Support (%)",
    sd_pct = "Standard Deviation"
  ) %>%
  fmt_number(
    columns = c(mean_pct, sd_pct),
    decimals = 2
  )

reliability_table
```

```{r}
#| label: fig-reliability
#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
#| echo: false
library(ggplot2)

# Create a more aesthetic bar chart with alpha and no labels
ggplot(summary_data, aes(x = reliability, y = mean_pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_pct - sd_pct, ymax = mean_pct + sd_pct), 
                width = 0.2, 
                position = position_dodge(0.7)) +
  labs(
    title = "Average Support Percentage by Pollster Reliability",
    subtitle = "Comparison of support percentages for Kamala Harris and Donald Trump",
    x = "Pollster Reliability",
    y = "Average Support (%)",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = "#1f78b4", "Donald Trump" = "#e31a1c")) + # Custom colors
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

```

### Impact of Methodology on Poll Results

Polling methodology is a critical factor that influences the results of any survey. Polls were categorized according to their methodology (e.g., Online Panel, Phone Interview) to evaluate how different polling methods might affect the percentage of support (pct) for each candidate. For each polling methodology, the average pct for Kamala Harris and Donald Trump was calculated and compared. This analysis aimed to determine whether specific methodologies consistently resulted in higher or lower support for either candidate. Comparison of candidate vote percentages by polling methodology is demonstrated by (@tbl-methodology) Discussion: We observed potential biases in support based on the polling method, providing insights into which methodologies might offer more reliable forecasts.

```{r}
#| label: tbl-methodology
#| tbl-cap: "Table of Methodologies comparing the Mean Support Percentage for Donald Trump and Kamala Harris with Difference (Kamala Harris - Donald Trump)"
#| echo: false

library(tidyverse)
library(gt)

# Define the 11 categories
categories <- c("Mail-to-Phone", "Mail-to-Web", "Email", "Live Phone", 
                "Online Ad", "Text", "IVR", "Probability Panel", 
                "Text-to-Web", "App Panel", "Online Panel")

# Split the methodology and categorize into the 11 categories
categorized_data <- data %>%
  filter(!is.na(methodology)) %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  separate_rows(methodology, sep = "/") %>%
  filter(methodology %in% categories) %>%
  group_by(candidate_name, methodology) %>%
  summarise(
    mean_pct = mean(pct, na.rm = TRUE),  # Calculate mean with NA removal if necessary
    .groups = 'drop'
  ) %>%
  pivot_wider(
    names_from = candidate_name,
    values_from = mean_pct,
    values_fill = list(`Donald Trump` = 0, `Kamala Harris` = 0)  # Fill missing values with 0
  )

# Rename the columns explicitly
categorized_data <- categorized_data %>%
  rename(
    `Donald Trump Mean Support (%)` = `Donald Trump`,
    `Kamala Harris Mean Support (%)` = `Kamala Harris`
  ) %>%
  mutate(
    Difference = `Kamala Harris Mean Support (%)` - `Donald Trump Mean Support (%)`
  )

# Create a table using the gt package
reliability_table <- categorized_data %>%
  gt() %>%
  tab_header(
    title = "Polling Data Summary: Methodology and Support Percentages",
    subtitle = "Mean Support Percentages (pct) for Kamala Harris and Donald Trump by Polling Methodology Category"
  ) %>%
  cols_label(
    methodology = "Polling Methodology",
    `Donald Trump Mean Support (%)` = "Trump(%)",
    `Kamala Harris Mean Support (%)` = "Harris(%)",
    Difference = "Difference"
  ) %>%
  fmt_number(
    columns = c(`Donald Trump Mean Support (%)`, `Kamala Harris Mean Support (%)`, Difference),
    decimals = 2
  )
# Display the table
reliability_table

```

### State-by-State Polling Trends

Given the importance of state-level polling data in U.S. presidential elections, this analysis focused on key battleground states (e.g., Florida, Pennsylvania, Wisconsin). The polling results for Kamala Harris and Donald Trump were examined by state (state), and the average pct was calculated for each swing state. These trends were visualized to assess whether one candidate consistently received more support in these critical states, as such patterns could significantly influence the final election outcome. A state-by-state breakdown of the polling percentages for each candidate, with a focus on swing states is shown by (@fig-state). Discussion: The results highlight key states where the election might be decided, helping to pinpoint potential trends favoring one candidate.

```{r}
#| label: fig-state
#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
#| echo: false
# Calculate average pct by state and candidate
library(tidyverse)
library(ggplot2)

# Filter out rows with NA values in the 'state' column and keep only rows for Kamala Harris and Donald Trump
filtered_data <- data %>%
  filter(!is.na(state)) %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))

# Calculate the average support percentage for each candidate by state
state_summary_data <- filtered_data %>%
  group_by(state, candidate_name) %>%
  summarise(
    avg_pct = mean(pct),
    .groups = 'drop'
  )

# Create a state-by-state breakdown bar plot of polling percentages
ggplot(state_summary_data, aes(x = state, y = avg_pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  labs(
    title = "State-by-State Polling Trends in Key Battleground States",
    subtitle = "Average support percentages for Kamala Harris and Donald Trump",
    x = "State",
    y = "Average Support Percentage (%)",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = "#1f78b4", "Donald Trump" = "#e31a1c")) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

```

### Sample Size and Poll Accuracy

The impact of sample size (sample_size) on the accuracy of polling results was examined. Larger sample sizes are typically considered more reliable, prompting the categorization of polls into small, medium, and large groups based on sample size. For each group, the average percentage of support (pct) for Kamala Harris and Donald Trump was calculated, and variability within each sample size category was analyzed to evaluate the reliability of the results. A chart(@fig-samplesize) comparing candidate vote percentages by sample size category. Discussion: The results suggest whether larger sample sizes produce more accurate and reliable forecasts, helping us to understand the potential limitations of smaller polls.

```{r}
#| label: fig-samplesize
#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
#| echo: false
# Categorizing sample size into small, medium, and large groups
library(tidyverse)
library(ggplot2)

# Remove rows with NA values in the 'sample_size' column and keep only rows for Kamala Harris and Donald Trump
filtered_data <- data %>%
  filter(!is.na(sample_size)) %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))

# Categorize sample size into small, medium, and large groups
filtered_data <- filtered_data %>%
  mutate(
    sample_size_category = case_when(
      sample_size <= 1000 ~ "Small",
      sample_size > 1000 & sample_size <= 1500 ~ "Medium",
      sample_size > 1500 ~ "Large"
    )
  )

# Calculate average pct and standard deviation by sample size category and candidate
summary_data_sample_size <- filtered_data %>%
  group_by(candidate_name, sample_size_category) %>%
  summarise(
    avg_pct = mean(pct),
    sd_pct = sd(pct),
    .groups = 'drop'
  )

# Create the chart comparing candidate vote percentages by sample size category
ggplot(summary_data_sample_size, aes(x = sample_size_category, y = avg_pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = avg_pct - sd_pct, ymax = avg_pct + sd_pct), 
                width = 0.2, 
                position = position_dodge(0.7)) +
  labs(
    title = "Candidate Vote Percentages by Sample Size Category",
    subtitle = "Average support percentages for Kamala Harris and Donald Trump based on sample size groups",
    x = "Sample Size Category",
    y = "Average Support Percentage (%)",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = "#1f78b4", "Donald Trump" = "#e31a1c")) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

```

Alternate Analysis Data 5: Sponsorship and Bias in Polling Finally, the potential bias of polls sponsored by partisan organizations was assessed to determine whether such sponsorship influenced support for the associated candidate. The vote percentages (pct) for Kamala Harris and Donald Trump were compared in polls sponsored by Democratic-leaning and Republican-leaning organizations (sponsor_candidate_party). By calculating the average pct for each candidate within partisan-sponsored polls, this analysis explored whether partisan sponsorship resulted in a systematic overestimation of support for a particular candidate. A comparison of the polling results in partisan-sponsored polls, broken down by party affiliation is shown by (@fig-sponsorship). Discussion: This analysis reveals potential biases in partisan-sponsored polls and assesses the objectivity of different polling organizations.

```{r}
#| label: fig-sponsorship
#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
#| echo: false
# Calculate the average pct and standard deviation for each candidate by sponsorship type
sponsorship_summary_data <- data %>%
  filter(sponsor_candidate_party != "Non-Partisan") %>%  # Focus on partisan-sponsored polls
  group_by(candidate_name, sponsor_candidate_party) %>%
  summarise(
    avg_pct = mean(pct),
    sd_pct = sd(pct),
    .groups = 'drop'
  )
library(ggplot2)

# Create the bar chart comparing vote percentages by sponsorship type
ggplot(sponsorship_summary_data, aes(x = sponsor_candidate_party, y = avg_pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = avg_pct - sd_pct, ymax = avg_pct + sd_pct), 
                width = 0.2, 
                position = position_dodge(0.7)) +
  labs(
    title = "Vote Percentages by Partisan Sponsorship",
    subtitle = "Average support percentages for Kamala Harris and Donald Trump in partisan-sponsored polls",
    x = "Sponsorship (Party Affiliation)",
    y = "Average Support Percentage (%)",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = "#1f78b4", "Donald Trump" = "#e31a1c")) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

```

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset. A thorough discussion of measurement, relating to the dataset, is provided in the data section. Please ensure that you explain how we went from some phenomena in the world that happened to an entry in the dataset that you are interested in.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

The model should be nicely written out, well-explained, justified, and appropriate.

Detail the statistical model used, defining and explaining each aspect and its importance, and ensure that variables are well-defined and correspond with those discussed in the data section.

The model should have an appropriate balance of complexity—neither overly simplistic nor unnecessarily complicated—and be justified as suitable for the situation.

## Modeling decisions

Explain how decisions made in modeling reflect the aspects discussed in the data section, including why specific features are included (e.g., why use age rather than age-groups, treating province effects as levels, categorizing gender, etc?).

## Mathematical notations

Present the model using appropriate mathematical notation supplemented with plain English explanations, defining every component. If applicable, define sensible priors for Bayesian models.

## Assumptions, limitations, circumstances

Clearly discuss the underlying assumptions, potential limitations, and circumstances where the model may not be appropriate.

## Software & model validation

Mention the software used to implement the model, and provide evidence of model validation and checking—such as out-of-sample testing, RMSE calculations, test/training splits, or sensitivity analyses—while addressing model convergence, diagnostics, and any alternative models or variants considered,

## Strenth and weeknesses and final choice

including their strengths and weaknesses and the rationale for the final model choice.

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

Results will likely require summary statistics, tables, graphs, images, and possibly statistical analysis or maps. There should also be text associated with all these aspects.

Show the reader the results by plotting them where possible. Talk about them. Explain them. That said, this section should strictly relay results.

Regression tables must not contain stars.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## What is done in this paper? {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## What is something that we learn about the world?

Please don't use these as sub-heading labels - change them to be what your point actually is.

## What is another thing that we learn about the world?

## Weaknesses and next steps

What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix 1 {.unnumbered}

YouGov has been a popular polling organization in predicting U.S. elections for a long time, they have experience predicting US elections in the past. One of the key reasons for us to select YouGov to dive into it is their innovative use of an advanced statistical model known as Multilevel Regression with Post-stratification (MRP). This model allows YouGov to have detailed predictions by using survey data with demographic background information, ensuring the representativeness of broader populations.

Moreover, FiveThirtyEight found that YouGov did not exhibit a strong house effect during the 2016 U.S. election, meaning they did not show a bias toward any political party(@n16), this further built up their credibility. Their model is intended to predict the votes of everyone in the US national voter file. YouGov’s MRP model used to predict support for each candidate in the US 2024 presidential election has three parts. Firstly, they will use people’s responses in the survey to estimate the likelihood of voting. From here, calculate their probability of voting for a specific party if people will vote. Combined and reweight responses, to give predictions for candidates (@br24).

## Frame and sample

YouGov's frame of the sample is people in the YouGov panel which use a nonprobability sampling. Everyone can join the YouGov panel, and YouGov will recruit participants (the sample) to the survey with a rigours process. YouGov recruits American adults through various advertising methods and partnerships. They also offer surveys in multiple languages.

By completing surveys, participants earn points which can be exchanged for monetary rewards or vendor equivalents (@m). YouGov will invite a representative set of panellists from their panel to take the survey. YouGov also have the resource to link participants to TargetSmart’s voter file, so they can ensure they are verified registered voters. They include precinct-level vote data of voters to make their sample more representative by improving the geographic representation of underrepresented areas (@br24). Other information used to invite survey participants who match the characteristics of the population of interest includes government data and other information collected from respondents when they join our panel like age, gender, race, education, etc (@m).

Most of the participants in their presidential election survey have decided to vote while a small proportion are undecided. Thus, the answer they get is only what people plan to do instead of committed actions. So, the survey result and built model are best to reflect the current stage of the race, not perfect for prediction in the future. People can even change their decision to vote or not over time, which adds more changes over time (@br24). This will affect YouGov’s model based on their methodology, and YouGov is reflecting these changes with an updated model corresponding.

## Survey and After-Survey Process

YouGov's surveys are all online, with any device and at any time the respondent prefers. Questions asked include who they will likely vote for, as well as their likelihood to vote in the actual election and other questions. The same almost sample is being used in each survey. In this way, participants are tracked over time, so YouGov can study their shifts as the campaign progresses. For example, they have discovered stability in voters’ candidate preferences for 2024 (@br24). YouGov ensures the reliability and validity of its surveys through high standards and transparent reporting. Participants’ privacy is protected by giving them control over the usage of their data such as allowing requests for data corrections and opting out of cookies. They aggregate responses in reporting to protect participants’ identities. Something good about their questionnaire is that participants will have the choice of “prefer not to say” and skipping the question (@m).

Another noticeable good point of their questionnaire is that it includes a broad range of topics including political, economic, environmental, and education issues, at the same time capturing people's potential engagement with voting by questions such as their likelihood of voting and method of voting. For the presidential election, YouGov has an unusually large sample compared to other opinion polling, with nearly 100,000 interviews in total contributing to their estimate model. They don’t only continue with this model; after September, they will update with re-interviewed responses from over 20,000 people (@br24). They build long-term relationships with their panellists.

For the presidential election, YouGov’s U.S. panellists have been surveyed regularly since December 2023, and these panellists do not only get interviewed once but instated, and engaged in monthly or quarterly re-interviews (@br24). But besides their effort to gather a large sample and actively include underrepresented regions by selective sampling, there are still areas that are hard to reach and lacking data groups that only have a small sample, such as people in small states and younger voters in larger states. Thus, after completing surveys, YouGov employs a weighting process to adjust respondents' influence based on their demographic characteristics and presidential vote (@br24). In helps to develop the inclusiveness of the data gathered,

## Quality of Data Gathered

To ensure accuracy, YouGov applies a strategy of monitoring, testing, and refinement. They have a team with techniques to catch unreliable respondents. They also apply consistency checks to people’s responses to avoid fraudulent responses. People who do not meet response quality standards will be removed from the final sample (@m). For example, in their poll from October 12-15, 1,869 started the survey initially. 145 were deleted due to break-offs, 100 more were removed for data quality purposes, reasons including short interview completion time, and failed attention check, failed consistent checks etc. So the reporting is based on the remaining 1,624 respondents.

## Weakness

Their surveys are completely online, and therefore limited to individuals with internet access. Also, there could be additional factors that could influence the results beyond the reported margin of error. These include how they are phrasing questions and respondent bias, all of which can introduce potential errors in the survey outcomes.

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: false
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: false
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-22 scripts/00-simulate_data.R

@@ -1,0 +1,93 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of US elections
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `Forcasting-US-Elections` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(lubridate)
+
+# Simulate a sample dataset
+set.seed(123)  # For reproducibility
+polling_data <- tibble(
+  poll_id = 1:10,
+  pollster = c("YouGov", "Ipsos", "SurveyUSA", "Morning Consult", "Quinnipiac", 
+               "Emerson College", "Marist College", "Rasmussen Reports", 
+               "Fox News", "CNN"),
+  sample_size = c(1500, 1200, 800, 1300, 1400, 1100, 1000, 1600, 900, 1250),
+  population = c("Likely Voters", "Registered Voters", "Likely Voters", 
+                 "Likely Voters", "Registered Voters", "Likely Voters", 
+                 "Registered Voters", "Likely Voters", "Likely Voters", 
+                 "Registered Voters"),
+  candidate_name = rep(c("Kamala Harris", "Donald Trump"), 5),
+  pct = c(46, 44, 47, 43, 49, 45, 48, 42, 47, 46)
+)
+
+sample_data <- tibble(
+  pollster_id = sample(1:100, 50, replace = TRUE),
+  pollster = sample(c("Pollster A", "Pollster B", "Pollster C"), 50, replace = TRUE),
+  sponsors = sample(c("Sponsor 1", "Sponsor 2", "Sponsor 3"), 50, replace = TRUE),
+  transparency_score = sample(c(NA, seq(0, 10, by = 0.5)), 50, replace = TRUE),
+  state = sample(c("CA", "TX", "NY", "FL", "PA"), 50, replace = TRUE),
+  start_date = sample(seq.Date(from = as.Date("2024-01-01"), to = as.Date("2024-10-01"), by = "day"), 50, replace = TRUE),
+  end_date = sample(seq.Date(from = as.Date("2024-01-02"), to = as.Date("2024-10-03"), by = "day"), 50, replace = TRUE),
+  sponsor_candidate_party = sample(c("Democratic", "Republican"), 50, replace = TRUE),
[…]

### 2024-10-22 scripts/03-clean_data.R

@@ -1,0 +1,57 @@
+#### Preamble ####
+# Purpose: Cleans the raw US election data
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read_csv("data/01-raw_data/president_polls.csv")
+
+# Assuming your data frame is called 'data'
+
+
+
+## filter out variables of interests
+filtered_data <- raw_data %>%
+  select(pollster_id, pollster, state,
+         start_date, end_date, sample_size, population, race_id,
+         cycle, election_date, candidate_name, pct, pollscore, methodology
+         , sponsor_candidate_party)
+filtered_data <- data %>%
+  filter(candidate_name %in% c("Donald Trump", "Kamala Harris"))
+column_types <- sapply(filtered_data, class)
+data.frame(Column = names(column_types), Type = column_types)
+
+# TODO: discuss whether to keep NA of states and sample size
+na_counts <- filtered_data %>%
+  summarise(across(everything(), ~ sum(is.na(.))))
+
+cleaned_data <- filtered_data %>%
+  mutate(
+    pollster_id = as.numeric(pollster_id), # Ensure pollster_id is numeric
+    sample_size = as.numeric(sample_size), # Ensure sample_size is numeric
+    race_id = as.numeric(race_id),         # Ensure race_id is numeric
+    cycle = as.numeric(cycle),             # Ensure cycle is numeric
+    pollscore = as.numeric(pollscore),     # Ensure pollscore is numeric
+    pct = as.numeric(pct),                 # Ensure pct is numeric
+    
+    # Convert character columns to lowercase for consistency and trim whitespace
+    pollster = str_trim(tolower(pollster)),
+    state = str_trim(tolower(state)),
+    population = str_trim(tolower(population)),
+    candidate_name = str_trim(tolower(candidate_name)),
+    methodology = str_trim(tolower(methodology)),
+    sponsor_candidate_party = str_trim(tolower(sponsor_candidate_party))
+  )
+
+# View the first few rows of the cleaned data (optional)
+head(cleaned_data)
+
+#### Save data ####
+write_csv(filtered_data, "data/02-analysis_data/analysis_data.csv")

### 2024-10-22 scripts/04-test_analysis_data.R

@@ -1,0 +1,113 @@
+#### Preamble ####
+# Purpose: Tests simulation and analysis data
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+
+data <- read_csv("data/02-analysis_data/analysis_data.csv")
+
+
+#### Test data ####
+# Test Case 1: Check for NAs in each variable
+na_counts <- sample_data %>%
+  summarise(across(everything(), ~ sum(is.na(.))))
+
+print("NA counts per variable:")
+print(na_counts)
+
+# Test Case 2: Check if start_date is earlier than end_date
+date_check <- sample_data %>%
+  mutate(valid_date = start_date <= end_date) %>%
+  summarise(all_dates_valid = all(valid_date))
+
+print("Are all start dates earlier than or equal to end dates?")
+print(date_check)
+
+# Test Case 3: Check if sample_size values are positive
+sample_size_check <- sample_data %>%
+  filter(!is.na(sample_size)) %>%
+  summarise(all_sample_sizes_positive = all(sample_size > 0))
+
+print("Are all sample sizes positive?")
+print(sample_size_check)
+
+# Test Case 4: Check if transparency_score is between 0 and 10
+transparency_check <- sample_data %>%
+  filter(!is.na(transparency_score)) %>%
+  summarise(valid_transparency_scores = all(transparency_score >= 0 & transparency_score <= 10))
+
+print("Are all transparency scores between 0 and 10?")
+print(transparency_check)
+
+# Test Case 5: Check if cycle is 2024 for all entries
+cycle_check <- sample_data %>%
+  summarise(all_cycles_2024 = all(cycle == 2024))
+
+print("Is the cycle 2024 for all entries?")
+print(cycle_check)
+
+
+
+
+
+# Test that the dataset has 151 rows - there are 151 divisions in Australia
+
+
+test_that("dataset has 151 rows", {
+  expect_equal(nrow(analysis_data), 151)
+})
+
+# Test that the dataset has 3 columns
+test_that("dataset has 3 columns", {
+  expect_equal(ncol(analysis_data), 3)
+})
+
+# Test that the 'division' column is character type
+test_that("'division' is character", {
[…]

### 2024-10-22 paper/paper.qmd
add citations to content and r used

@@
 presidential election is almost at its final stage, with all candidates actively campaigning across numerous states.
-The presidential election will happen on November 5, 2024 (@calendar).
+The presidential election will happen on November 5, 2024 [@calendar].
 The two major parties, Democrats and Republicans, are in particular focus.
 Former President Donald Trump represents the Republican party, and President Joe Biden initially led re-election for the Democrats.
-However, Biden has dropped out, and Kamala Harris has taken over as the Democratic nominee (@g24), who is the first Black woman nominee.
+However, Biden has dropped out, and Kamala Harris has taken over as the Democratic nominee [@g24], who is the first Black woman nominee.
 Harris is the Both candidates are focusing their efforts on securing votes by addressing their aspect on social issues.
@@
 The polling outcomes allow campaigns of different parties to strategically target specific demographics.
-For example in 2012, pollsters and data modellers such as Nate Silver have all used survey research and statistical models to successfully predict the result of Barack Obama's victory in the presidential election with consistent and reliable forecasts (@b17).
+For example in 2012, pollsters and data modellers such as Nate Silver have all used survey research and statistical models to successfully predict the result of Barack Obama's victory in the presidential election with consistent and reliable forecasts [@b17].
 
@@
 It provides a comprehensive view of voter preferences through aggregated results from numerous national polls conducted by various polling organizations.
-The dataset was simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2].
+The dataset was simulated, cleaned, analyzed, and tested using the R programming language [@talia], tidyverse [@citetidyverse], knitr [@citeknitr], ggplot2 [@citeggplot2], gt[@gt].
 
@@
 These variables allow us to explore various dimensions of polling data, such as trends in voter support across different states and the influence of pollster reliability on polling outcomes.
-The (@tbl-firstten) below provides a snapshot of the polling data, displaying the first ten entries.
+The [@tbl-firstten] below provides a snapshot of the polling data, displaying the first ten entries.
 This includes the pollster names, sample sizes, and the percentage of support for each candidate, which can reveal patterns in public opinion and help identify how different organizations may report varying levels of support for Kamala Harris and Donald Trump.
@@
 Mean pct values were calculated for each category, and standard deviation was included to illustrate variability within the poll results.
-The table(@tbl-reliability) and bar chart(@fig-reliability) showing the average `pct` for each candidate across different levels of pollster reliability.
+The table[@tbl-reliability] and bar chart[@fig-reliability] showing the average `pct` for each candidate across different levels of pollster reliability.
 Discussion: This analysis highlights whether pollsters with higher reliability scores offer more accurate predictions and if their estimates favor one candidate over the other.
@@
 This analysis aimed to determine whether specific methodologies consistently resulted in higher or lower support for either candidate.
-Comparison of candidate vote percentages by polling methodology is demonstrated by (@tbl-methodology) Discussion: We observed potential biases in support based on the polling method, providing insights into which methodologies might offer more reliable forecasts.
+Comparison of candidate vote percentages by polling methodology is demonstrated by [@tbl-methodology] Discussion: We observed potential biases in support based on the polling method, providing insights into which methodologies might offer more reliable forecasts.
 
@@
 These trends were visualized to assess whether one candidate consistently received more support in these critical states, as such patterns could significantly influence the final election outcome.
-A state-by-state breakdown of the polling percentages for each candidate, with a focus on swing states is shown by (@fig-state).
+A state-by-state breakdown of the polling percentages for each candidate, with a focus on swing states is shown by [@fig-state].
 Discussion: The results highlight key states where the election might be decided, helping to pinpoint potential trends favoring one candidate.
@@
 For each group, the average percentage of support (pct) for Kamala Harris and Donald Trump was calculated, and variability within each sample size category was analyzed to evaluate the reliability of the results.
-A chart(@fig-samplesize) comparing candidate vote percentages by sample size category.
+A chart[@fig-samplesize] comparing candidate vote percentages by sample size category.
 Discussion: The results suggest whether larger sample sizes produce more accurate and reliable forecasts, helping us to understand the potential limitations of smaller polls.
@@
 By calculating the average pct for each candidate within partisan-sponsored polls, this analysis explored whether partisan sponsorship resulted in a systematic overestimation of support for a particular candidate.
-A comparison of the polling results in partisan-sponsored polls, broken down by party affiliation is shown by (@fig-sponsorship).
+A comparison of the polling results in partisan-sponsored polls, broken down by party affiliation is shown by [@fig-sponsorship].
 Discussion: This analysis reveals potential biases in partisan-sponsored polls and assesses the objectivity of different polling organizations.
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Some of our data is of penguins [@fig-bills], from @palmerpenguins.
 
@@
 
-And also planes (@fig-planes).
+And also planes [@fig-planes].
 (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
@@
 ```
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@talia] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 
@@
 Moreover, FiveThirtyEight found that YouGov did not exhibit a strong house effect during the 2016 U.S.
-election, meaning they did not show a bias toward any political party(@n16), this further built up their credibility.
+election, meaning they did not show a bias toward any political party[@n16], this further built up their credibility.
 Their model is intended to predict the votes of everyone in the US national voter file.
@@
 From here, calculate their probability of voting for a specific party if people will vote.
-Combined and reweight responses, to give predictions for candidates (@br24).
+Combined and reweight responses, to give predictions for candidates [@br24].
 
@@
 
-By completing surveys, participants earn points which can be exchanged for monetary rewards or vendor equivalents (@m).
+By completing surveys, participants earn points which can be exchanged for monetary rewards or vendor equivalents [@m].
 YouGov will invite a representative set of panellists from their panel to take the survey.
 YouGov also have the resource to link participants to TargetSmart’s voter file, so they can ensure they are verified registered voters.
-They include precinct-level vote data of voters to make their sample more representative by improving the geographic representation of underrepresented areas (@br24).
-Other information used to invite survey participants who match the characteristics of the population of interest includes government data and other information collected from respondents when they join our panel like age, gender, race, education, etc (@m).
+They include precinct-level vote data of voters to make their sample more representative by improving the geographic representation of underrepresented areas [@br24].
+Other information used to invite survey participants who match the characteristics of the population of interest includes government data and other information collected from respondents when they join our panel like age, gender, race, education, etc [@m].
 
@@
 So, the survey result and built model are best to reflect the current stage of the race, not perfect for prediction in the future.
-People can even change their decision to vote or not over time, which adds more changes over time (@br24).
+People can even change their decision to vote or not over time, which adds more changes over time [@br24].
 This will affect YouGov’s model based on their methodology, and YouGov is reflecting these changes with an updated model corresponding.
@@
 In this way, participants are tracked over time, so YouGov can study their shifts as the campaign progresses.
-For example, they have discovered stability in voters’ candidate preferences for 2024 (@br24).
+For example, they have discovered stability in voters’ candidate preferences for 2024 [@br24].
 YouGov ensures the reliability and validity of its surveys through high standards and transparent reporting.
@@
 They aggregate responses in reporting to protect participants’ identities.
-Something good about their questionnaire is that participants will have the choice of “prefer not to say” and skipping the question (@m).
+Something good about their questionnaire is that participants will have the choice of “prefer not to say” and skipping the question [@m].
 
@@
 For the presidential election, YouGov has an unusually large sample compared to other opinion polling, with nearly 100,000 interviews in total contributing to their estimate model.
-They don’t only continue with this model; after September, they will update with re-interviewed responses from over 20,000 people (@br24).
+They don’t only continue with this model; after September, they will update with re-interviewed responses from over 20,000 people [@br24].
 They build long-term relationships with their panellists.
@@
 For the presidential election, YouGov’s U.S.
-panellists have been surveyed regularly since December 2023, and these panellists do not only get interviewed once but instated, and engaged in monthly or quarterly re-interviews (@br24).
+panellists have been surveyed regularly since December 2023, and these panellists do not only get interviewed once but instated, and engaged in monthly or quarterly re-interviews [@br24].
 But besides their effort to gather a large sample and actively include underrepresented regions by selective sampling, there are still areas that are hard to reach and lacking data groups that only have a small sample, such as people in small states and younger voters in larger states.
-Thus, after completing surveys, YouGov employs a weighting process to adjust respondents' influence based on their demographic characteristics and presidential vote (@br24).
+Thus, after completing surveys, YouGov employs a weighting process to adjust respondents' influence based on their demographic characteristics and presidential vote [@br24].
 In helps to develop the inclusiveness of the data gathered,
@@
 They also apply consistency checks to people’s responses to avoid fraudulent responses.
-People who do not meet response quality standards will be removed from the final sample (@m).
+People who do not meet response quality standards will be removed from the final sample [@m].
 For example, in their poll from October 12-15, 1,869 started the survey initially.

### 2024-10-31 scripts/03-clean_data.R
update clean data and model

@@ -1,38 +1,54 @@
 #### Preamble ####
-# Purpose: Cleans the raw US election data
+# Purpose: Clean the variables so that the data focuses on donald trump and the variables of interest for further analysis and model building
 # Author: Amy Jin
-# Date: 18 October 2024
-# Contact: amyzh.jin@mail.utoronto,ca
+# Date: 27 October 2024
+# Contact: amyzh.jin@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: raw data in the correct directory and corret file name that has the required varibles
+# Assumption: These varibles are workable for data analysis and contributes to model building
 
-#### Workspace setup ####
+#### Workspace setup and loading data  ####
+
 library(tidyverse)
+library(janitor)
+library(arrow)#for parquet
 
-#### Clean data ####
-raw_data <- read_csv("data/01-raw_data/president_polls.csv")
-
-# Assuming your data frame is called 'data'
-
+# Load the raw polling data
+raw_data <- read_csv("data/01-raw_data/president_polls.csv")|>
+  clean_names()
 
 
-## filter out variables of interests
-filtered_data <- raw_data %>%
+#### Data Cleaning ####
+raw_data <- raw_data %>%
+  mutate(is_national = if_else(is.na(state), 1, 0))  # Create binary for national polls
+
+
+# Filter data to Harris estimates based on high-quality polls after she declared
+filtered_data <- raw_data |>
+  filter(
+    candidate_name == "Donald Trump",
+    numeric_grade >= 2.5 
+    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  mutate(
+    num_trump = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+  )
+
+
+# Select relevant columns for analysis
+filtered_data <- filtered_data %>%
   select(pollster_id, pollster, state,
          start_date, end_date, sample_size, population, race_id,
-         cycle, election_date, candidate_name, pct, pollscore, methodology
[…]

### 2024-10-31 scripts/06-model_data.R

@@ -1,37 +1,130 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Create, compare, and select the best model to predict the 2024 US election result
+# Author: [Your Name]
+# Date: [Update with today's date]
+# Contact: [Your Email]
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Ensure all necessary data and libraries are available
+# Any other information needed? [Update with additional context]
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(broom)
+library(ggplot2)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+#### Define models ####
 
+# Model 1: pct as a function of end_date (time trend only)
+model_date <- lm(pct ~ end_date, data = analysis_data)
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+# Model 2: pct as a function of end_date and pollster (time trend + pollster effect)
+model_date_pollster <- lm(pct ~ end_date + pollster, data = analysis_data)
+
+# Model 3: pct as a function of end_date, pollster, and is_national (includes national vs. state poll effect)
+model_date_pollster_national <- lm(pct ~ end_date + pollster + is_national, data = analysis_data)
+
+# Model 4: Bayesian model using rstanarm
+model_bayesian <- stan_glm(pct ~ end_date + pollster + is_national,
+                           data = analysis_data,
+                           family = gaussian,
+                           prior = normal(0, 5),
+                           chains = 4, iter = 2000)
+
+#### Compare models ####
+# Calculate AIC and BIC for model comparison
[…]

### 2024-11-01 paper/paper.qmd
update paper and citations

@@
 ---
-title: "Forcasting US Election 2024"
+title: "Forecasting the US Election 2024"
 subtitle: "##TODO: result here"
@@
 Harris is the Both candidates are focusing their efforts on securing votes by addressing their aspect on social issues.
-pollsters and election forecasting with data evidence are important in predicting election outcomes.
+
+Pollsters and election forecasting with data evidence are important in predicting election outcomes.
 Polling can highlight underlying or emerging issues, as well as reveal voter preferences.
@@
 
-## Overview
+## Overview {#sec-dataoverview}
 
@@
 
-The dataset comprises several key variables of interest, including poll_id, a unique identifier for each poll conducted; pollster, which indicates the organization conducting the poll; sample_size, representing the total number of respondents; population, specifying the voting group described (e.g., likely voters); candidate_name, the names of the candidates in the poll (e.g., Kamala Harris, Donald Trump); and pct, the percentage of the vote or support received by each candidate.
+The dataset comprises several key variables of interest, including but not limited to,
+
+**poll_id:** a unique identifier for each poll conducted pollster: which indicates the organization conducting the poll
+
+**sample_size:** representing the total number of respondents
+
+**population:** specifying the voting group described (e.g., likely voters)
+
+**candidate_name:** the names of the candidates in the poll (e.g., Kamala Harris, Donald Trump)
+
+**pct:** the percentage of the vote or support received by each candidate.
+
 These variables allow us to explore various dimensions of polling data, such as trends in voter support across different states and the influence of pollster reliability on polling outcomes.
-The [@tbl-firstten] below provides a snapshot of the polling data, displaying the first ten entries.
+The table below provides a snapshot of the polling data, displaying the first ten entries.
 This includes the pollster names, sample sizes, and the percentage of support for each candidate, which can reveal patterns in public opinion and help identify how different organizations may report varying levels of support for Kamala Harris and Donald Trump.
+
+### Response variables:
+
+The main response variable of our analysis is the percentage support (pct) for Donald Trump and Kamala Harris shown in each poll.
+This number tells us what percentage of people said they would vote for each candidate when the survey was taken.
+We use it as the basis of our electoral prediction as it helps us understand how people feel about the candidates.
+
+TODO: descibe
+
+```{r}
+#| label: tbl-pct
+#| tbl-cap: "Histogram of Trump Support Percentage (pct)"
+#| echo: false
+ggplot(data, aes(x = pct)) +
+  geom_histogram(binwidth = 5, fill = "#e31a1c", color = "black") +
+  theme_minimal() +
+  labs(
+    x = "Percentage Support (pct)",
+    y = "Frequency"
+  )
+```
+
+### Predictor variables:
+
+Our model takes into account various variables; their names and descriptions are,
+
+**Pollster**: The name of the polling organization that conducted the poll (e.g., YouGov, RMG Research).
+
+**State:** The U.S. state where the poll was conducted.
+
+**Sample Size:** The total number of respondents that participated in the poll. End Date: The date the poll ended.
+
+These variables were selected and kept due to their evident relationship with the response variable and the significance of it.
+More details of the relationships and model are provided in later sections.
 
@@
 #| label: tbl-firstten
-#| tbl-cap: "Sample sizes and support percentages for Kamala Harris and Donald Trump"
+#| tbl-cap: "Polling Data Snapshot: 2024 U.S. Presidential Election"
 #| echo: false
@@
   gt() %>%
-  tab_header(
-    title = "Polling Data Snapshot: 2024 U.S. Presidential Election",
-  ) %>%
   cols_label(
@@
 
+### Data Manipulation and cleaning:
+
+The data was cleaned, filtered and mutated to better align with our prediction model.
+The software and packages used have been highlighted in @sec-dataoverview.
+The key steps of our cleaning process were:
+
+Filtered out lower quality polls by only keeping polls that met a set numeric grade threshold (numeric_grade \>= 2.5).
+Filtered out pollsters that had less than 5 entries.
+Filtered out data that was collected before July 21st, 2024.
+Added a variable num_trump to count the number of votes for Donald Trump using pct.
+For data entries where the state was missing, they were categorised as “National”, to standardize the data.
+
+The dataset had several other variables that we did not include in our analysis. These variables were either majorly missing values or not informative enough to be included.
+
 ## Summary statistics & Relationships
@@
 Mean pct values were calculated for each category, and standard deviation was included to illustrate variability within the poll results.
-The table[@tbl-reliability] and bar chart[@fig-reliability] showing the average `pct` for each candidate across different levels of pollster reliability.
+The table (@tbl-reliability) and bar chart @fig-reliability showing the average `pct` for each candidate across different levels of pollster reliability.
 Discussion: This analysis highlights whether pollsters with higher reliability scores offer more accurate predictions and if their estimates favor one candidate over the other.
@@
 #| label: tbl-reliability
-#| tbl-cap: "Mean and Standard Deviation of Support Percentages (pct) for Kamala Harris and Donald Trump Across Pollster Reliability Levels"
+#| tbl-cap: "Summary Statistics for Support Percentages Across Pollster Reliability Levels"
 #| echo: false
@@
   gt() %>%
-  tab_header(
-    title = "Polling Data Summary: Pollster Reliability and Support Percentages"
-  ) %>%
   cols_label(
@@
 #| label: fig-reliability
-#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
+#| fig-cap: "Average Support Percentage by Pollster Reliability"
 #| echo: false
@@
   labs(
-    title = "Average Support Percentage by Pollster Reliability",
     subtitle = "Comparison of support percentages for Kamala Harris and Donald Trump",
@@
   gt() %>%
-  tab_header(
-    title = "Polling Data Summary: Methodology and Support Percentages",
-    subtitle = "Mean Support Percentages (pct) for Kamala Harris and Donald Trump by Polling Methodology Category"
-  ) %>%
   cols_label(
@@
 
-### State-by-State Polling Trends
-
-Given the importance of state-level polling data in U.S.
-presidential elections, this analysis focused on key battleground states (e.g., Florida, Pennsylvania, Wisconsin).
-The polling results for Kamala Harris and Donald Trump were examined by state (state), and the average pct was calculated for each swing state.
-These trends were visualized to assess whether one candidate consistently received more support in these critical states, as such patterns could significantly influence the final election outcome.
-A state-by-state breakdown of the polling percentages for each candidate, with a focus on swing states is shown by [@fig-state].
-Discussion: The results highlight key states where the election might be decided, helping to pinpoint potential trends favoring one candidate.
-
-```{r}
-#| label: fig-state
-#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
-#| echo: false
-# Calculate average pct by state and candidate
-library(tidyverse)
-library(ggplot2)
-
-# Filter out rows with NA values in the 'state' column and keep only rows for Kamala Harris and Donald Trump
-filtered_data <- data %>%
-  filter(!is.na(state)) %>%
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))
-
-# Calculate the average support percentage for each candidate by state
-state_summary_data <- filtered_data %>%
-  group_by(state, candidate_name) %>%
-  summarise(
-    avg_pct = mean(pct),
-    .groups = 'drop'
-  )
-
-# Create a state-by-state breakdown bar plot of polling percentages
-ggplot(state_summary_data, aes(x = state, y = avg_pct, fill = candidate_name)) +
-  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
-  labs(
-    title = "State-by-State Polling Trends in Key Battleground States",
-    subtitle = "Average support percentages for Kamala Harris and Donald Trump",
-    x = "State",
-    y = "Average Support Percentage (%)",
-    fill = "Candidate"
-  ) +
-  scale_fill_manual(values = c("Kamala Harris" = "#1f78b4", "Donald Trump" = "#e31a1c")) +
-  theme_minimal(base_size = 14) +
-  theme(
-    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
-    plot.subtitle = element_text(hjust = 0.5, size = 12),
-    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
-    axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    legend.position = "top",
-    legend.title = element_text(face = "bold")
-  )
-
-```
-
 ### Sample Size and Poll Accuracy
@@
 #| label: fig-samplesize
-#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
+#| fig-cap: "Candidate Vote Percentages by Sample Size Category"
 #| echo: false
@@
   labs(
-    title = "Candidate Vote Percentages by Sample Size Category",
-    subtitle = "Average support percentages for Kamala Harris and Donald Trump based on sample size groups",
     x = "Sample Size Category",
@@
 
-Alternate Analysis Data 5: Sponsorship and Bias in Polling Finally, the potential bias of polls sponsored by partisan organizations was assessed to determine whether such sponsorship influenced support for the associated candidate.
+### Sponsorship and Bias in Polling
+
+Finally, the potential bias of polls sponsored by partisan organizations was assessed to determine whether such sponsorship influenced support for the associated candidate.
 The vote percentages (pct) for Kamala Harris and Donald Trump were compared in polls sponsored by Democratic-leaning and Republican-leaning organizations (sponsor_candidate_party).
@@
 #| label: fig-sponsorship
-#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
+#| fig-cap: "Vote Percentages by Partisan Sponsorship"
 #| echo: false
@@
   labs(
-    title = "Vote Percentages by Partisan Sponsorship",
-    subtitle = "Average support percentages for Kamala Harris and Donald Trump in partisan-sponsored polls",
     x = "Sponsorship (Party Affiliation)",
@@
 
-## Measurement
+## Measurement and Limitations
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
-A thorough discussion of measurement, relating to the dataset, is provided in the data section.
-Please ensure that you explain how we went from some phenomena in the world that happened to an entry in the dataset that you are interested in.
+This dataset comes with several important measurement and limitations:
 
-## Outcome variables
+**Polling Accuracy:** While efforts are made to prioritize high-quality polls through numeric ratings, differences in polling methods can still introduce some degree of bias.
+Pollster ratings enhance reliability but cannot fully eliminate methodological inconsistencies.
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+**Time Sensitivity:** The dataset captures voter preferences at particular points rather than over time, providing only a static view.
+This limitation means that shifts in voter sentiment throughout the campaign are not represented.
 
-Some of our data is of penguins [@fig-bills], from @palmerpenguins.
+**Geographical Coverage:** The dataset includes both national and state-level polling, but coverage is uneven.
+High-interest states, like swing states, are polled more frequently, while data from other regions may be sparse, potentially affecting regional analysis.
 
-Talk more about it.
+**Participation Bias**: Even with strict rules for polling, certain biases remain possible, such as selection bias in survey participation and response bias due to social desirability factors.
 
-And also planes [@fig-planes].
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Statistical Model and Analysis
 
-Talk way more about it.
+## Model Overview
 
-## Predictor variables
+The goal of our modeling is to predict which candidate from Donald Trump and Kamala Harris will win the 2024 US election.
+We will firstly, predict support for the two candidates using the tailored cleaned data in the 2024 US presidential election using aggregated polling data, and secondly, to select a model that balances complexity and interpretability while accurately capturing the key trends and effects.
+This section describes the statistical models used, justifies the modeling choices, and discusses the assumptions, limitations, and validation strategies.
+Background details, including diagnostics and additional model exploration, are available in [Appendix -@sec-model-details].
 
-Add graphs, tables and text.
+We utilized both traditional linear models and Bayesian models.
+The Bayesian approach, in particular, allows us to incorporate prior beliefs and better quantify uncertainty, which is crucial given the inherent variability in polling data.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+#### Model Specifications
 
-# Model
+We constructed four models of increasing complexity:
 
-The goal of our modelling strategy is twofold. Firstly,...
+**Model 1**: A linear regression where the dependent variable,
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+pct, the percentage of respondents supporting each candidate, is modeled as a function of the date the poll ended,
 
-The model should be nicely written out, well-explained, justified, and appropriate.
+end_date, to capture temporal trends.
 
-Detail the statistical model used, defining and explaining each aspect and its importance, and ensure that variables are well-defined and correspond with those discussed in the data section.
+**Model 2**: An extension of Model 1 that includes a categorical variable,
 
-The model should have an appropriate balance of complexity—neither overly simplistic nor unnecessarily complicated—and be justified as suitable for the situation.
+pollster, to account for systematic differences between polling organizations.
 
-## Modeling decisions
+**Model 3**: Adds a binary indicator,
 
-Explain how decisions made in modeling reflect the aspects discussed in the data section, including why specific features are included (e.g., why use age rather than age-groups, treating province effects as levels, categorizing gender, etc?).
+is_national, to Model 2 to differentiate between national and state-specific polls.
 
-## Mathematical notations
+**Model 4**: A Bayesian model implemented using the `rstanarm` package, incorporating the same predictors as Model 3 but allowing for more robust estimation under uncertainty.
 
-Present the model using appropriate mathematical notation supplemented with plain English explanations, defining every component. If applicable, define sensible priors for Bayesian models.
+### We mathematically define these models as follows:
 
-## Assumptions, limitations, circumstances
+#### Model 1
 
-Clearly discuss the underlying assumptions, potential limitations, and circumstances where the model may not be appropriate.
+$$
+y_i = \beta_0 + \beta_1 \cdot \text{end\_date}_i + \epsilon_i, \quad \epsilon_i \sim \text{Normal}(0, \sigma^2)
+$$
 
-## Software & model validation
+#### Model 2
 
-Mention the software used to implement the model, and provide evidence of model validation and checking—such as out-of-sample testing, RMSE calculations, test/training splits, or sensitivity analyses—while addressing model convergence, diagnostics, and any alternative models or variants considered,
+$$
+y_i = \beta_0 + \beta_1 \cdot \text{end\_date}_i + \sum_{j=1}^J \gamma_j \cdot \text{pollster}_{ij} + \epsilon_i, \quad \epsilon_i \sim \text{Normal}(0, \sigma^2)
+$$
 
-## Strenth and weeknesses and final choice
+#### Model 3
 
-including their strengths and weaknesses and the rationale for the final model choice.
+$$
+y_i = \beta_0 + \beta_1 \cdot \text{end\_date}_i + \sum_{j=1}^J \gamma_j \cdot \text{pollster}_{ij} + \delta \cdot \text{is\_national}_i + \epsilon_i, \quad \epsilon_i \sim \text{Normal}(0, \sigma^2)
+$$
 
-## Model set-up
+#### Model 4 (Bayesian)
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+$$
+y_i \mid \mu_i, \sigma \sim \text{Normal}(\mu_i, \sigma), \quad \mu_i = \beta_0 + \beta_1 \cdot \text{end\_date}_i + \sum_{j=1}^J \gamma_j \cdot \text{pollster}_{ij} + \delta \cdot \text{is\_national}_i
+$$
 
-```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-```
-We run the model in R [@talia] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+**Priors for Model 4:**
 
-### Model justification
+$$
+\beta_0, \beta_1, \gamma_j, \delta \sim \text{Normal}(0, 5), \quad \sigma \sim \text{Exponential}(1)
+$$
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+#### Variables and Justification
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+1.  **Dependent Variable**: ( y_i ) (support percentage, `pct`).
+2.  **Predictors**:
+    -   **(** \text{end\_date} ): Captures time trends, which are essential in understanding shifts in public opinion.
+    -   **(** \text{pollster} ): Accounts for differences in polling methodologies and potential biases.
+    -   **(** \text{is\_national} ): Differentiates between polls conducted at the national level and those specific to individual states, as national polls tend to exhibit different characteristics and sample compositions compared to state polls.
+
+These modeling choices reflect insights from the data section, where we observed systematic variations in support percentages over time, across pollsters, and between national and state polls.
+
+#### Model Justification
+
+The modeling approach was designed to balance interpretability and predictive power:
+
+**Model 1** serves as a baseline, capturing the temporal trend.
+
+**Model 2** addresses variability between pollsters, which is crucial given differences in methodologies and sample recruitment.
+
+**Model 3** adds another layer of nuance by accounting for the national vs. state-specific distinction.
+
+**Model 4** employs Bayesian inference, allowing us to incorporate uncertainty and use prior distributions for more robust predictions.
+
+The Bayesian approach is particularly useful given the variability in polling data.
+Using `rstanarm`, we employ default weakly informative priors, which are appropriate given the lack of strong prior information.
+These priors help regularize the estimates and prevent overfitting.
+
+#### Assumptions, Limitations, and Circumstances
+
+1.  **Assumptions**:
+    -   Linear relationships between predictors and the outcome.
+    -   Independence of observations, which may be violated if polls from the same organization are correlated.
+    -   Normality of residuals in linear models.
+2.  **Limitations**:
+    -   The models do not account for potential interactions between variables or non-linear effects, which could improve predictive accuracy.
+    -   The Bayesian model's convergence can be sensitive to the choice of priors and the data's variability.
+3.  **Circumstances**:
+    -   The models may be less appropriate in highly volatile electoral environments where sudden shifts in public opinion occur.
+    -   The temporal model assumes a relatively smooth trend, which may not hold in the presence of significant events influencing voter sentiment.
+
+#### Software and Model Validation
+
+The models were implemented in R using the `tidyverse` and `rstanarm` packages. We assessed model performance using:
+
+**AIC/BIC**: For model selection based on goodness-of-fit and penalization for complexity.
+
+**Cross-validation**: (To be detailed further in the appendix) to evaluate out-of-sample performance.
+
+**Residual Diagnostics**: To check for violations of model assumptions.
+
+**Model Convergence**: Monitored using diagnostics provided by `rstanarm` for the Bayesian model.
+
+#### Strengths, Weaknesses, and Final Model Choice
+
+-   **Strengths**: The models are interpretable and straightforward, with the Bayesian approach providing a robust framework for uncertainty quantification.
+-   **Weaknesses**: Potential oversimplification and assumptions of linearity. Future iterations could explore non-linear effects or hierarchical models.
+-   **Final Choice**: We selected **Model 3** (linear model with `end_date`, `pollster`, and `is_national`) for its balance of interpretability and predictive performance.
+The Bayesian model serves as a robustness check, with results presented in the appendix.
+
+Our final model selection and analysis provide a reliable and transparent prediction framework for Kamala Harris’s support in the 2024 US election.
 
@@
 
+\newpage
+
+# Appendix B: Idealized Survey Methodology and Survey Design
+
+## Methodology Overview
+
+We will be developing a prediction model and conducting a survey targeting eligible voters in the United States for the 2024 presidential election.
+The study will observe key demographic groups influencing voting behaviour, such as age, ethnicity, income, education, and political affiliation [@leighley2013who].
+Employing a combined stratified and quota sampling approach, we will ensure representation across these demographics, with a particular emphasis on swing states to enhance the accuracy of our predictions [@fitzgerald2024seven].
+Recruitment will utilize both online and offline methods, including community outreach and random digit dialling, while incentives will be offered to boost participation.
+Data validation will involve screening questions and consistency checks to ensure reliable responses.
+Finally, we will implement hybrid modelling techniques to aggregate survey data and apply statistical models for more accurate electoral predictions, correcting for any biases through weighting methods [@pasek2015predicting] [@wlezien2002timeline].
+A copy of our designed survey will be attached.
+
+## Target Population
+
+Target Population: The target population for our prediction model and survey includes all eligible voters across the United States participating in the upcoming 2024 presidential election.
+Population context study: We will observe specific key groups within the population that will affect people’s voting actions.
+These include age groups, racial and ethnic, educational levels geographic regions, political affiliation, etc.
+[@leighley2013who].
+
+## Sampling Approach
+
+We will combine Stratified Sampling and Quota Sampling.
+Stratification: Stratified sampling involves dividing the population into distinct subgroups (strata) and then randomly sampling from each stratum (Thompson, 2012).
+This will ensure that each demographic group is represented in the sample, allowing for a more informative analysis of voter behaviour within each group.
+
+### Quota Sampling within Strata:
+
+To improve representation, quota sampling is used within each stratum.
+This will involve collecting data until the predetermined quotas for each stratum are met (Thompson, 2012).
+This approach will ensure that the sample is more representative of the broader population, better to make a generalization.
+
+### Idefitiying Strata:
+
+The target population consists of eligible voters in the U.S., which can be further divided into strata based on key demographic characteristics and grouping factors identified in the previous stage.
+The criteria for defining these strata include age, ethnicity, income level, education, and geographic location [@leighley2013who].
+For example, we can have: Age: 18-29, 30-44, 45-59, 60+ Income: Low-income (\<\$40,000), Middle-income (\$40,000-\$100,000), High-income (\>\$100,000) Political Affiliation: Democrat, Republican, Independent, Other Set Quotas: Within the defined strata, we will establish quotas for each subgroup to ensure adequate representation based on key demographic features.
+These quotas will be determined using recent census data and voter registration records, in line with methodologies from reliable national polling (Keeter et al., 2016).
+A well-designed quota system will enhance balanced representation across the defined strata.
+For example, we can implement: Age: 18-29: 15% (180 respondents) 30-44: 25% (300 respondents) 45-59: 30% (360 respondents) 60+: 30% (360 respondents) Swing States: An important planning consideration is our focus on swing states to enhance the precision and efficiency of our election forecast.
+Swing states, characterized by competitive races, significantly influence the final outcome.
+In the 2024 election, these states could realistically be won by either Democrat Kamala Harris or Republican Donald Trump.
+Key swing states include Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, where both major campaigns are actively targeting undecided voters [@fitzgerald2024seven].
+
+To achieve this, we will:
+
+Increase Quota Allocation for Swing States: Allocate a higher portion of the sample size to respondents from swing states to ensure these areas are more heavily represented.
+Adjust Stratification and Quotas by Region: Implement customized quota settings within each swing state to capture its unique demographics.
+This approach allows us to predict the voting patterns of distinct groups, increasing the likelihood of capturing accurate outcomes.
+For example: In Michigan, we will allocate more resources and further divide our efforts by ethnicity.
+The state has a significant Arab-American population, and the attitudes of both major parties toward this group will likely influence their decisions, which can greatly impact the overall outcome of the state [@fitzgerald2024seven].
+
+In contrast, states like Hawaii will be excluded from our analysis, as it has consistently leaned Democratic since its admission in 1959, indicating a solid Democratic stronghold [@thehill2024hawaii].
+
+### Respondent Recruitment
+
+#### Online Recruitment Methods
+
+With advancements in technology, online recruitment has become increasingly cost-efficient.
+This approach will direct potential respondents to an online survey platform where they can easily participate in the study.
+
+Email Outreach: Collaborations with community organizations and advocacy groups will facilitate email outreach to their members.
+
+Online Panels: Utilization of established online survey panels will possibly make recruitment efficient.
+
+Social Media Advertising: Targeted advertisements will be utilized on platforms such as Facebook, Twitter, and Instagram to reach specific demographic groups based on age, location, and political interests.
+While this approach effectively captures the desired participant groups, we must remain cautious of potential sampling bias.
+
+#### Offline Recruitment Methods
+
+Phone Outreach - Random Digit Dialing (RDD): We will employ a random-digit dialling method to recruit respondents for phone interviews.
+This approach allows us to reach individuals who may not be engaged online, particularly older demographics and those in rural areas.
+However, this traditional recruitment format will constitute only a small portion of our overall efforts, as advancements in technology have made online recruitment significantly more efficient.
+We plan to use phone outreach as a coplelement when online recruitment does not yield sufficient participants.
+This approach helps avoid inclusion errors by reaching populations without internet access or those who are not engaged in online communities.
+However, we face challenges such as increasing non-response rates due to the public's growing reluctance to participate in phone surveys and the rising use of technologies that screen unsolicited calls [@wang2015forecasting].
+
+### Incentives for Participation
+
+Singer and Ye [@singer2013use] discuss incentives in surveys can significantly enhance response rates and reduce nonresponse bias, leading to more accurate data collection.
+However, it is essential to strike a balance to avoid low-quality data resulting from participants motivated solely by the incentive; excessively large rewards can skew results.
+Monetary Compensation: Respondents will have the opportunity to receive a monetary reward of approximately \$20 for doing the survey.
+Gift Cards: Alternatively, we can offer gift cards to popular retailers or online platforms as incentives.
+Additionally, we may collaborate with funding resources willing to provide funds, enabling us to distribute gift cards while simultaneously promoting their brand.
+
+### Extra Effort on Specific Demographics:
+
+We will keep monitoring the demographic composition of respondents throughout the data collection process to ensure that specific groups meet their quotas.
+If certain demographics are underrepresented, we will implement targeted outreach efforts, such as collaborating with organizations that can help raise awareness of the survey within those communities.
+This approach will enhance participation from these groups.
+If the collected sample still does not align with the desired proportions indicated by the population, we will perform post-stratification adjustments (weighting) to correct any imbalances.
+
+### Data Validation
+
+Screening Questions: We will implement initial screening questions to ensure the data we collect reflects reality.
+This will include verifying participants' voting eligibility by confirming their citizenship and age.
+Consistency Checks: Drawing from YouGov's questionnaire methodology, we will include multiple questions in various formats to assess a single point, such as voting intention.
+By identifying inconsistent responses in this manner, we can enhance the accuracy of our data.
+Participants with heavily inconsistent answers will be noted, and if discrepancies are significant, their responses will be excluded from the final analysis.
+
+### Poll Aggregation and Modeling with Hybrid Approaches
+
+Adopting Hybrid Models for Election Forecasting: According to Pasek [@pasek2015predicting], surveys have become essential tools for forecasting outcomes with three primary strategies for pooling survey data: aggregation, predictive modelling, and hybrid models.
+From his study, the hybrid approach has the most potential to accurately pool election information and make predictions.
+It combines both aggregation and prediction.
+Here, we will be using a traditional hybrid model, which aggregates survey data and uses statistical models to predict.
+
+#### Survey Aggregation:
+
+This involves pooling results from multiple surveys, weekly rolling averages will be used to smooth out random fluctuations and limit random error by reducing the uncertainty of the estimates [@pasek2015predicting].
+
+#### Predictive Modeling:
+
+Our models might include variables such as demographic information, past election results, and economic indicators.
+The goal is to build a model that use statistical techniques to analyze these factors and identify patterns which help to forecast potential electoral outcomes [@pasek2015predicting].
+Some specific poll-based forecasting models can be used to mitigate the tendency for survey results to overreact to campaign events by accounting for how polls are expected to relate to voter behaviour and discounting temporary shifts [@wlezien2002timeline].
+Integration: The most effective aspect of our hybrid model is its ability to integrate both aggregated survey data and results from predictive models.
+This allows us to leverage individual survey results alongside predictive elements such as trends and patterns, ultimately refining our predictive outcomes.
+By doing so, the model not only generates meaningful predictions but also enhances our understanding of the interplay among various electoral forces [@pasek2015predicting].
+
+#### Weighting:
+
+Despite our best efforts to ensure the sample accurately reflects the population during the sampling stage, non-response bias and responses being discarded during consistency checks may still result in an imperfect representation of the population.
+Similar to methodologies used by YouGov, we will apply weights to adjust responses based on demographic representation within the voting population.
+This approach helps correct for any over- or under-sampled groups.
+
+#### Trade-off:
+
+By implementing a complex hybrid model that requires careful weighting adjustments, we are able to ensure that each demographic group's influence aligns with its actual proportion in reality; making the conclusion more easily generated by the population.
+At the same time, as with all complex models, this brings more challenges in data analysis, making it hard to interpret the result.
+This also requires more time and financial investment
+
+### Survey Questions and Design
+
+#### Online surveys:
+
+We will be using online surveys because of the advantages offered by technology and expanding internet access across regions.
+Online surveys provide flexibility and convenience, enabling us to customize questions to fit specific contexts, order, and coverage of questions.
+It is also cost-effective, especially when reaching large, geographically dispersed samples, as in our case.
+Additionally, they facilitate real-time data monitoring to ensure data quality, while reducing the administrative and logistical costs associated with traditional survey methods [@evans2005value].
+
+#### Question Structure:
+
+The survey design will be designed to ensure all questions remain with clarity, neutrality, and inclusivity.
+which helps avoid introducing systematic bias.
+We will avoid leading language that could skew responses predictably.
+By reducing ambiguity, we enhance response quality and reliability, thereby reducing measurement error and ensuring our data represents a broad spectrum of opinions.
+Question Type: The survey will primarily use Likert scales and multiple-choice formats to capture insights on voting intentions and demographics, both crucial for understanding potential election outcomes.
+These are decided to gather high-quality responses with high usability in prediction.
+
+#### Demographic:
+
+To accurately reflect real-world diversity, the survey will include demographic questions, such as those on race, ethnicity, and education level, with an inclusive range of options.
+This minimizes sampling bias by offering choices that accommodate diverse identities, thus helping participants feel represented and reducing the oversimplification of our sample.
+For example, in the gender question, we’ll include options like “Woman,” “Man,” “Transgender,” “Non-binary/non-conforming,” “Prefer not to answer,” and “Other” (with a text box to specify).
+This approach allows participants to self-identify accurately, reducing both sampling bias and response variance across different groups.
+
+#### Voting intention:
+
+This section will assess participants' likelihood of voting, designed to minimize social desirability bias, a type of response bias where the participant might change their answer to appear more socially responsible and maintain themselves in better self-pictures [@grimm2010social].
+We will be using a Likert scale to enable nuanced responses without implying a “correct” answer.
+This way,with less biased data, we can capture a range of genuine attitudes (such as: “On a scale of 1–5, how likely are you to vote in the upcoming election?”)
+
+#### Political Preference:
+
+In this section, we delve into participants’ existing preferences for specific political parties.
+To reduce variance from inconsistent answers, we will ask multiple questions about party preference, including questions on past voting behaviour.
+By phrasing similar questions in varied ways, we can confirm consistency in responses (such as: “Which political party do you currently support?” and “Which party did you vote for in the last election?”)
+
+#### Social Issues:
+
+This section aims to explore participants' viewpoints on influential issues in this presidential election, such as economic policy, healthcare, and immigration [@nadeem2024issues].
+Questions will be designed free from framing bias, this is where specific wording might unduly influence responses [@entman2007framing].
+Neutral wording helps capture participants' genuine opinions on issues most likely to impact their voting behaviour.
+
+### Budget Allocation Breakdown: Total of \$100K
+
+Survey Design and Implementation (30% - \$30,000): Questionnaire development, pilot testing, and data collection.
+Sampling and Recruitment (25% - \$25,000): Stratified sampling, participant incentives, and outreach.
+Data Validation and Cleaning (15% - \$15,000): Screening and data preparation.
+Data Analysis and Modeling (20% - \$20,000): Software licenses, model development, and bias correction.
+Reporting and Dissemination (10% - \$10,000): Report and present to the public.
+Contingency Fund (5% - \$5,000): For unexpected costs.
+
+### Google Forms Survey Link
+
+Include a link to your Google Forms survey here. Survey Link Copy of the Survey Questions
+
 # Additional data details

### 2024-11-02 paper/paper.qmd
model section complete

@@
 library(tidyr)
+# Load the arrow package
+library(arrow)
+library(here)
 data = read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# Read the Parquet file
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+
 ```
@@
 presidential election between Kamala Harris and Donald Trump.The dataset was obtained from https://projects.fivethirtyeight.com/polls/president-general/2024/national/ or (FiveThirtyEight’s “Poll of Polls” for the 2024 U.S.
-Presidential election [@fivethirtyeight]).
+Presidential election [@fivethirtyeight2024]).
 It provides a comprehensive view of voter preferences through aggregated results from numerous national polls conducted by various polling organizations.
@@
 
+The results and details are in @sec-model-details.
+
 #### Strengths, Weaknesses, and Final Model Choice
@@
 
-## Posterior predictive check
+## Model Validation and Evaluation
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+Our goal in developing predictive models for the 2024 US presidential election is to ensure that our models are not only accurate but also robust and generalizable.
+To achieve this, we implement a comprehensive model validation process.
+This section describes our approach to model validation, including criteria used to compare models, diagnostic techniques employed to evaluate model assumptions, and specific methods to ensure the reliability of our Bayesian model.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+We validate our models using several strategies: Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) for model selection, cross-validation to evaluate out-of-sample performance, and diagnostic plots to assess model assumptions.
+Additionally, we perform posterior predictive checks and convergence diagnostics for the Bayesian model.
+
+#### Software and Implementation
+The models were implemented in R using the `tidyverse` and `rstanarm` packages.
+The validation process was executed using tools from these packages and additional packages, such as `broom` and `loo`, for model diagnostics.
+
+### Model Validation Process
+
+#### 1. AIC and BIC for Model Selection
+To compare the linear models' goodness-of-fit while penalizing complexity, we use AIC and BIC.
+These metrics help identify the most appropriate model for the data, balancing explanatory power and parsimony.
+The results are presented in (@tbl-AIC):
+
+```{r, echo=FALSE, include=FALSE, warning=FALSE}
+library(tidyverse)
+library(broom)
+library(rstanarm)
+library(ggplot2)
+library(loo)
+
+# Define the models
+model_date <- lm(pct ~ end_date, data = analysis_data)
+model_date_pollster <- lm(pct ~ end_date + pollster, data = analysis_data)
+model_date_pollster_national <- lm(pct ~ end_date + pollster + is_national, data = analysis_data)
+model_bayesian <- stan_glm(
+  pct ~ end_date + pollster + is_national,
+  data = analysis_data,
+  family = gaussian,
+  prior = normal(0, 5),
+  chains = 4, iter = 2000
+)
+```
 
 ```{r}
-#| eval: false
+#| label: tbl-AIC
+#| tbl-cap: "Model Comparison Table: AIC and BIC Values for Model Selection"
 #| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+# Load required librarie
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+# Create the model comparison table using gt
+model_comparison <- tibble(
+  Model = c("model_date", "model_date_pollster", "model_date_pollster_national"),
+  AIC = c(AIC(model_date), AIC(model_date_pollster), AIC(model_date_pollster_national)),
+  BIC = c(BIC(model_date), BIC(model_date_pollster), BIC(model_date_pollster_national))
+)
 
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+# Print the model comparison table using gt
+model_comparison %>%
+  gt() %>%
+  fmt_number(
+    columns = c(AIC, BIC),
+    decimals = 2
+  )
+
 ```
 
-## Diagnostics
+**Results**: Model 3 has the lowest AIC and BIC, suggesting it offers the best balance between goodness-of-fit and complexity.
+This model includes both time trends and categorical effects for pollsters and national versus state-specific polls.
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+#### 2. Cross-Validation for Out-of-Sample Performance
+We use leave-one-out cross-validation (LOO) for the Bayesian model to assess predictive performance.
+The `loo` package computes the expected log predictive density (ELPD), which indicates how well the model generalizes to new data.
+Table (@tbl-CV) shows the result.
 
 ```{r}
+#| label: tbl-CV
+#| tbl-cap: "LOO Summary for Bayesian Model LOOCV Result"
 #| echo: false
-#| eval: false
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
 
-plot(first_model, "trace")
+# Load necessary libraries
+library(gt)
+library(loo)
 
-plot(first_model, "rhat")
+# Run LOO on the Bayesian model
+loo_result <- loo(model_bayesian)
+
+# Extract summary statistics from loo_result
+loo_summary <- tibble(
+  Metric = c("LOOIC", "SE(LOOIC)", "p_loo", "elpd_loo", "SE(elpd_loo)"),
+  Value = c(
+    loo_result$estimates["looic", "Estimate"],
+    loo_result$estimates["looic", "SE"],
+    loo_result$estimates["p_loo", "Estimate"],
+    loo_result$estimates["elpd_loo", "Estimate"],
+    loo_result$estimates["elpd_loo", "SE"]
+  )
+)
+
+# Create a gt table for the LOO summary
+loo_summary %>%
+  gt() %>%
+  fmt_number(
+    columns = "Value",
+    decimals = 2
+  )
+
 ```
+
+**Cross-Validation Results**:
+The LOO estimate for the Bayesian model shows that the model has good predictive performance, supporting its use for forecasting election outcomes.
+
+#### 3. Residual Diagnostics for Linear Models
+We generate residual plots to check for violations of assumptions such as homoscedasticity (constant variance of residuals) and linearity.
+
+```{r, echo=FALSE}
+ggplot(augment(model_date), aes(x = .fitted, y = .resid)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  theme_minimal() +
+  labs(title = "Residual Plot for Model 1", x = "Fitted Values", y = "Residuals")
+
+# Model 2 Residual Plot
+ggplot(augment(model_date_pollster), aes(x = .fitted, y = .resid)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  theme_minimal() +
+  labs(title = "Residual Plot for Model 2", x = "Fitted Values", y = "Residuals")
+
+# Model 3 Residual Plot
+ggplot(augment(model_date_pollster_national), aes(x = .fitted, y = .resid)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  theme_minimal() +
+  labs(title = "Residual Plot for Model 3", x = "Fitted Values", y = "Residuals")
+```
+
+- **Model 1 (Simple Time Trend)**: The residuals show a pattern, suggesting potential model misfit.
+- **Model 2 (Time Trend + Pollster)**: The residuals are more evenly spread, indicating an improved model fit.
+- **Model 3 (Time Trend + Pollster + National/State)**: The residuals show no obvious patterns, suggesting that the model assumptions are reasonably met.
+
+**Interpretation**: Model 3's residuals suggest that the model adequately captures the data's underlying structure, with no major violations of the linear model assumptions.
+
+#### 4. Bayesian Model Diagnostics
+For the Bayesian model, we perform the following diagnostics:
+
+##### a. Posterior Predictive Check
+We use `pp_check()` from `rstanarm` to visualize how well the model's predictions align with the observed data.
+
+```{r, echo=FALSE}
+pp_check(model_bayesian) +
+  theme_classic() +
+  theme(legend.position = "bottom") +
+  labs(title = "Posterior Predictive Check for Bayesian Model")
+```
+
+**Interpretation**: The predictive distribution closely matches the observed data, indicating a good model fit.
+
+##### b. Convergence Diagnostics
+We assess MCMC convergence using trace plots and Rhat values:
+- **Trace Plots**: Show that the chains mix well and converge to a common distribution, indicating no issues with convergence.
+- **Rhat Values**: All Rhat values are close to 1, further confirming convergence.
+
+```{r, echo=FALSE}
+plot(model_bayesian, "trace")
+plot(model_bayesian, "rhat")
+
+```
+
+**Conclusion**: The Bayesian model's diagnostics confirm that the model converges well, and the posterior samples are reliable.
 

### 2024-11-02 scripts/00-simulate_data.R
r scripts

@@ -13,22 +13,24 @@ library(tidyverse)
 library(dplyr)
 library(lubridate)
 
-# Simulate a sample dataset
-set.seed(123)  # For reproducibility
+library(tibble)
+
+# Set the seed for reproducibility
+set.seed(106)
+
+# Simulate polling data
 polling_data <- tibble(
   poll_id = 1:10,
   pollster = c("YouGov", "Ipsos", "SurveyUSA", "Morning Consult", "Quinnipiac", 
                "Emerson College", "Marist College", "Rasmussen Reports", 
                "Fox News", "CNN"),
-  sample_size = c(1500, 1200, 800, 1300, 1400, 1100, 1000, 1600, 900, 1250),
-  population = c("Likely Voters", "Registered Voters", "Likely Voters", 
-                 "Likely Voters", "Registered Voters", "Likely Voters", 
-                 "Registered Voters", "Likely Voters", "Likely Voters", 
-                 "Registered Voters"),
+  sample_size = sample(800:1600, 10, replace = TRUE),
+  population = sample(c("Likely Voters", "Registered Voters"), 10, replace = TRUE),
   candidate_name = rep(c("Kamala Harris", "Donald Trump"), 5),
-  pct = c(46, 44, 47, 43, 49, 45, 48, 42, 47, 46)
+  pct = sample(40:60, 10, replace = TRUE)
 )
 
+# Simulate sample data
 sample_data <- tibble(
   pollster_id = sample(1:100, 50, replace = TRUE),
   pollster = sample(c("Pollster A", "Pollster B", "Pollster C"), 50, replace = TRUE),
@@ -48,46 +50,7 @@ sample_data <- tibble(
   candidate_id = sample(100:200, 50, replace = TRUE)
 )
 
-# View the simulated data
-head(sample_data)
-
-
-set.seed(853)
-
-
-#### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
-)
-
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
[…]

### 2024-11-02 scripts/01-test_simulated_data.R

@@ -1,89 +1,92 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
+# Purpose: Tests the structure and validity of the simulated polling dataset
 # Author: Rohan Alexander
 # Date: 26 September 2024
 # Contact: rohan.alexander@utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
+# - The `tidyverse` package must be installed and loaded
+# - 00-simulate_data.R must have been run
 # Any other information needed? Make sure you are in the `starter_folder` rproj
-
 
 #### Workspace setup ####
 library(tidyverse)
 
+# Load the simulated dataset
 analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
 
-# Test if the data was successfully loaded
+# Test if the dataset was successfully loaded
 if (exists("analysis_data")) {
   message("Test Passed: The dataset was successfully loaded.")
 } else {
   stop("Test Failed: The dataset could not be loaded.")
 }
 
-
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# 1. Check if the dataset has 50 rows
+if (nrow(analysis_data) == 50) {
+  message("Test Passed: The dataset has 50 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 50 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# 2. Check if the dataset has 15 columns
+if (ncol(analysis_data) == 16) {
+  message("Test Passed: The dataset has 16 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 16 columns.")
 }
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
+# 3. Check if 'pollster_id' contains only numeric values
+if (all(is.numeric(analysis_data$pollster_id))) {
+  message("Test Passed: 'pollster_id' contains only numeric values.")
[…]

### 2024-11-02 scripts/03-clean_data.R

@@ -195,36 +195,7 @@ filtered_data <- filtered_data %>%
   group_by(pollster) %>%
   filter(n() >= 5) %>%
   ungroup()
-#### Plot data ####
-base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Trump percent", x = "Date")
 
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point() +
-  geom_smooth()
-
-# Color by pollster
-# This gets messy - need to add a filter - see line 21
-base_plot +
-  geom_point(aes(color = pollster)) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-# Facet by pollster
-# Make the line 21 issue obvious
-# Also - is there duplication???? Need to go back and check
-base_plot +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(vars(pollster))
-
-# Color by pollscore
-base_plot +
-  geom_point(aes(color = factor(pollscore))) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
 
 #### Save cleaned data ####
 

### 2024-11-02 scripts/04-test_analysis_data.R

@@ -11,103 +11,81 @@
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(arrow)
+library(here)
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
+#### Preamble ####
+# Purpose: Tests simulation and analysis data
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: tidyverse, testthat, arrow, and here packages installed
+# Any other information needed? Make sure the data file is in the correct directory
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+library(arrow)
+library(here)
+
+# Load the data
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 
 #### Test data ####
-# Test Case 1: Check for NAs in each variable
-na_counts <- sample_data %>%
-  summarise(across(everything(), ~ sum(is.na(.))))
 
-print("NA counts per variable:")
-print(na_counts)
-
-# Test Case 2: Check if start_date is earlier than end_date
-date_check <- sample_data %>%
-  mutate(valid_date = start_date <= end_date) %>%
-  summarise(all_dates_valid = all(valid_date))
-
-print("Are all start dates earlier than or equal to end dates?")
-print(date_check)
-
-# Test Case 3: Check if sample_size values are positive
-sample_size_check <- sample_data %>%
-  filter(!is.na(sample_size)) %>%
-  summarise(all_sample_sizes_positive = all(sample_size > 0))
-
-print("Are all sample sizes positive?")
-print(sample_size_check)
-
-# Test Case 4: Check if transparency_score is between 0 and 10
-transparency_check <- sample_data %>%
-  filter(!is.na(transparency_score)) %>%
-  summarise(valid_transparency_scores = all(transparency_score >= 0 & transparency_score <= 10))
-
-print("Are all transparency scores between 0 and 10?")
-print(transparency_check)
-
-# Test Case 5: Check if cycle is 2024 for all entries
-cycle_check <- sample_data %>%
-  summarise(all_cycles_2024 = all(cycle == 2024))
-
-print("Is the cycle 2024 for all entries?")
-print(cycle_check)
-
-
-
-
-
[…]

### 2024-11-02 scripts/00-simulate_data.R
Delete scripts/00-simulate_data.R

@@ -1,56 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates a dataset of US elections
-# Author: Amy Jin
-# Date: 18 October 2024
-# Contact: amyzh.jin@mail.utoronto,ca
-# License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `Forcasting-US-Elections` rproj
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(dplyr)
-library(lubridate)
-
-library(tibble)
-
-# Set the seed for reproducibility
-set.seed(106)
-
-# Simulate polling data
-polling_data <- tibble(
-  poll_id = 1:10,
-  pollster = c("YouGov", "Ipsos", "SurveyUSA", "Morning Consult", "Quinnipiac", 
-               "Emerson College", "Marist College", "Rasmussen Reports", 
-               "Fox News", "CNN"),
-  sample_size = sample(800:1600, 10, replace = TRUE),
-  population = sample(c("Likely Voters", "Registered Voters"), 10, replace = TRUE),
-  candidate_name = rep(c("Kamala Harris", "Donald Trump"), 5),
-  pct = sample(40:60, 10, replace = TRUE)
-)
-
-# Simulate sample data
-sample_data <- tibble(
-  pollster_id = sample(1:100, 50, replace = TRUE),
-  pollster = sample(c("Pollster A", "Pollster B", "Pollster C"), 50, replace = TRUE),
-  sponsors = sample(c("Sponsor 1", "Sponsor 2", "Sponsor 3"), 50, replace = TRUE),
-  transparency_score = sample(c(NA, seq(0, 10, by = 0.5)), 50, replace = TRUE),
-  state = sample(c("CA", "TX", "NY", "FL", "PA"), 50, replace = TRUE),
-  start_date = sample(seq.Date(from = as.Date("2024-01-01"), to = as.Date("2024-10-01"), by = "day"), 50, replace = TRUE),
-  end_date = sample(seq.Date(from = as.Date("2024-01-02"), to = as.Date("2024-10-03"), by = "day"), 50, replace = TRUE),
-  sponsor_candidate_party = sample(c("Democratic", "Republican"), 50, replace = TRUE),
-  endorsed_candidate_name = sample(c("Candidate A", "Candidate B"), 50, replace = TRUE),
[…]

### 2024-11-02 scripts/00-simulate_data.R
simulation

@@ -1,0 +1,56 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of US elections
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `Forcasting-US-Elections` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(lubridate)
+
+library(tibble)
+
+# Set the seed for reproducibility
+set.seed(106)
+
+# Simulate polling data
+polling_data <- tibble(
+  poll_id = 1:10,
+  pollster = c("YouGov", "Ipsos", "SurveyUSA", "Morning Consult", "Quinnipiac", 
+               "Emerson College", "Marist College", "Rasmussen Reports", 
+               "Fox News", "CNN"),
+  sample_size = sample(800:1600, 10, replace = TRUE),
+  population = sample(c("Likely Voters", "Registered Voters"), 10, replace = TRUE),
+  candidate_name = rep(c("Kamala Harris", "Donald Trump"), 5),
+  pct = sample(40:60, 10, replace = TRUE)
+)
+
+# Simulate sample data
+sample_data <- tibble(
+  pollster_id = sample(1:100, 50, replace = TRUE),
+  pollster = sample(c("Pollster A", "Pollster B", "Pollster C"), 50, replace = TRUE),
+  sponsors = sample(c("Sponsor 1", "Sponsor 2", "Sponsor 3"), 50, replace = TRUE),
+  transparency_score = sample(c(NA, seq(0, 10, by = 0.5)), 50, replace = TRUE),
+  state = sample(c("CA", "TX", "NY", "FL", "PA"), 50, replace = TRUE),
+  start_date = sample(seq.Date(from = as.Date("2024-01-01"), to = as.Date("2024-10-01"), by = "day"), 50, replace = TRUE),
+  end_date = sample(seq.Date(from = as.Date("2024-01-02"), to = as.Date("2024-10-03"), by = "day"), 50, replace = TRUE),
+  sponsor_candidate_party = sample(c("Democratic", "Republican"), 50, replace = TRUE),
+  endorsed_candidate_name = sample(c("Candidate A", "Candidate B"), 50, replace = TRUE),
[…]

### 2024-11-02 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,18 +1,21 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of US elections
+# Purpose: Simulates a dataset of US elections with sample polling data
 # Author: Amy Jin
 # Date: 18 October 2024
-# Contact: amyzh.jin@mail.utoronto,ca
+# Contact: amyzh.jin@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `Forcasting-US-Elections` rproj
+# Pre-requisites: 
+#   - The `tidyverse` package must be installed and loaded
+#   - Ensure that you are in the `Forcasting-US-Elections` R project directory
+#   - The directory "data/00-simulated_data/" must exist for saving the data file
+# Any other information needed? 
+#   - Run this script to generate and save the simulated dataset
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(dplyr)
 library(lubridate)
-
 library(tibble)
 
 # Set the seed for reproducibility
@@ -49,8 +52,6 @@ sample_data <- tibble(
   election_date = rep(as.Date("2024-11-05"), 50),
   candidate_id = sample(100:200, 50, replace = TRUE)
 )
-
-
 
 #### Save data ####
 write_csv(sample_data, "data/00-simulated_data/simulated_data.csv")

### 2024-11-02 scripts/00-simulate_data.R
Delete scripts directory

@@ -1,57 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates a dataset of US elections with sample polling data
-# Author: Amy Jin
-# Date: 18 October 2024
-# Contact: amyzh.jin@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-#   - The `tidyverse` package must be installed and loaded
-#   - Ensure that you are in the `Forcasting-US-Elections` R project directory
-#   - The directory "data/00-simulated_data/" must exist for saving the data file
-# Any other information needed? 
-#   - Run this script to generate and save the simulated dataset
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(dplyr)
-library(lubridate)
-library(tibble)
-
-# Set the seed for reproducibility
-set.seed(106)
-
-# Simulate polling data
-polling_data <- tibble(
-  poll_id = 1:10,
-  pollster = c("YouGov", "Ipsos", "SurveyUSA", "Morning Consult", "Quinnipiac", 
-               "Emerson College", "Marist College", "Rasmussen Reports", 
-               "Fox News", "CNN"),
-  sample_size = sample(800:1600, 10, replace = TRUE),
-  population = sample(c("Likely Voters", "Registered Voters"), 10, replace = TRUE),
-  candidate_name = rep(c("Kamala Harris", "Donald Trump"), 5),
-  pct = sample(40:60, 10, replace = TRUE)
-)
-
-# Simulate sample data
-sample_data <- tibble(
-  pollster_id = sample(1:100, 50, replace = TRUE),
-  pollster = sample(c("Pollster A", "Pollster B", "Pollster C"), 50, replace = TRUE),
-  sponsors = sample(c("Sponsor 1", "Sponsor 2", "Sponsor 3"), 50, replace = TRUE),
-  transparency_score = sample(c(NA, seq(0, 10, by = 0.5)), 50, replace = TRUE),
-  state = sample(c("CA", "TX", "NY", "FL", "PA"), 50, replace = TRUE),
-  start_date = sample(seq.Date(from = as.Date("2024-01-01"), to = as.Date("2024-10-01"), by = "day"), 50, replace = TRUE),
-  end_date = sample(seq.Date(from = as.Date("2024-01-02"), to = as.Date("2024-10-03"), by = "day"), 50, replace = TRUE),
-  sponsor_candidate_party = sample(c("Democratic", "Republican"), 50, replace = TRUE),
[…]

### 2024-11-02 scripts/01-test_simulated_data.R

@@ -1,92 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the structure and validity of the simulated polling dataset
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
-# Pre-requisites: 
-# - The `tidyverse` package must be installed and loaded
-# - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
-#### Workspace setup ####
-library(tidyverse)
-
-# Load the simulated dataset
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-
-# Test if the dataset was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
-
-#### Test data ####
-
-# 1. Check if the dataset has 50 rows
-if (nrow(analysis_data) == 50) {
-  message("Test Passed: The dataset has 50 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 50 rows.")
-}
-
-# 2. Check if the dataset has 15 columns
-if (ncol(analysis_data) == 16) {
-  message("Test Passed: The dataset has 16 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 16 columns.")
-}
-
-# 3. Check if 'pollster_id' contains only numeric values
-if (all(is.numeric(analysis_data$pollster_id))) {
-  message("Test Passed: 'pollster_id' contains only numeric values.")
-} else {
-  stop("Test Failed: 'pollster_id' contains non-numeric values.")
-}
-
-# 4. Check if 'sample_size' values are within the expected range (500 to 2000)
-if (all(na.omit(analysis_data$sample_size) >= 500 & na.omit(analysis_data$sample_size) <= 2000)) {
-  message("Test Passed: 'sample_size' values are within the expected range.")
-} else {
-  stop("Test Failed: 'sample_size' values are outside the expected range.")
-}
-
-
-# 6. Check if 'population' contains only valid categories
-valid_populations <- c("Adults", "Likely Voters", "Registered Voters")
-if (all(analysis_data$population %in% valid_populations)) {
-  message("Test Passed: 'population' contains only valid categories.")
-} else {
[…]

### 2024-11-02 scripts/03-clean_data.R

@@ -1,203 +1,0 @@
-#### Preamble ####
-# Purpose: Clean the variables so that the data focuses on donald trump and the variables of interest for further analysis and model building
-# Author: Amy Jin
-# Date: 27 October 2024
-# Contact: amyzh.jin@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: raw data in the correct directory and corret file name that has the required varibles
-# Assumption: These varibles are workable for data analysis and contributes to model building
-
-#### Workspace setup and loading data  ####
-
-library(tidyverse)
-library(janitor)
-library(arrow)#for parquet
-
-# Load the raw polling data
-raw_data <- read_csv("data/01-raw_data/president_polls.csv")|>
-  clean_names()
-
-
-#### Data Cleaning ####
-raw_data <- raw_data %>%
-  mutate(is_national = if_else(is.na(state), 1, 0))  # Create binary for national polls
-
-
-# Filter data to Harris estimates based on high-quality polls after she declared
-filtered_data <- raw_data |>
-  filter(
-    candidate_name == "Donald Trump",
-    numeric_grade >= 2.5 
-    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
-  ) |>
-  mutate(
-    state = if_else(is.na(state), "National", state),
-    end_date = mdy(end_date)
-  ) |>
-  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
-  mutate(
-    num_trump = round((pct / 100) * sample_size, 0) # Need number not percent for some models
-  )
-
-
-# Select relevant columns for analysis
-filtered_data <- filtered_data %>%
-  select(pollster_id, pollster, state,
-         start_date, end_date, sample_size, population, race_id,
-         cycle, election_date, candidate_name, pct, numeric_grade, methodology
-         , num_trump, is_national)
-na_counts <- filtered_data %>%
-  summarise(across(everything(), ~ sum(is.na(.))))
-filtered_data <- filtered_data %>%
-  mutate(
-    pollster_id = as.numeric(pollster_id), # Ensure pollster_id is numeric
-    sample_size = as.numeric(sample_size), # Ensure sample_size is numeric
-    race_id = as.numeric(race_id),         # Ensure race_id is numeric
-    cycle = as.numeric(cycle),             # Ensure cycle is numeric
[…]

### 2024-11-02 scripts/04-test_analysis_data.R

@@ -1,91 +1,0 @@
-#### Preamble ####
-# Purpose: Tests simulation and analysis data
-# Author: Amy Jin
-# Date: 18 October 2024
-# Contact: amyzh.jin@mail.utoronto,ca
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(testthat)
-library(arrow)
-library(here)
-analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
-
-
-#### Preamble ####
-# Purpose: Tests simulation and analysis data
-# Author: Amy Jin
-# Date: 18 October 2024
-# Contact: amyzh.jin@mail.utoronto,ca
-# License: MIT
-# Pre-requisites: tidyverse, testthat, arrow, and here packages installed
-# Any other information needed? Make sure the data file is in the correct directory
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(testthat)
-library(arrow)
-library(here)
-
-# Load the data
-analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
-
-#### Test data ####
-
-# 1. Test if the data was loaded successfully
-test_that("Dataset loads correctly", {
-  expect_true(exists("analysis_data"))
-})
-
-# 2. Test if the data has rows
-test_that("Dataset has rows", {
-  expect_gt(nrow(analysis_data), 0)
-})
-
-# 3. Test if the data has columns
-test_that("Dataset has columns", {
-  expect_gt(ncol(analysis_data), 0)
-})
-
-# 4. Test if there are no missing values in essential columns
-test_that("No missing values in essential columns", {
-  essential_cols <- c("pollster_id", "pollster", "state", "start_date", "end_date", "sample_size")
-  expect_true(all(!is.na(analysis_data[ , essential_cols])))
-})
-
-# 5. Test if `sample_size` is always greater than zero
-test_that("Sample size is positive", {
-  expect_true(all(analysis_data$sample_size > 0))
-})
-
-# 6. Test if `start_date` is before `end_date`
-test_that("Start date is before end date", {
-  expect_true(all(as.Date(analysis_data$start_date) < as.Date(analysis_data$end_date)))
-})
-
-# 7. Test if `pct` values are between 0 and 100
-test_that("Percentages are between 0 and 100", {
[…]

### 2024-11-02 scripts/06-model_data.R

@@ -1,130 +1,0 @@
-#### Preamble ####
-# Purpose: Create, compare, and select the best model to predict the 2024 US election result
-# Author: [Your Name]
-# Date: [Update with today's date]
-# Contact: [Your Email]
-# License: MIT
-# Pre-requisites: Ensure all necessary data and libraries are available
-# Any other information needed? [Update with additional context]
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-library(broom)
-library(ggplot2)
-
-#### Read data ####
-analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
-
-#### Define models ####
-
-# Model 1: pct as a function of end_date (time trend only)
-model_date <- lm(pct ~ end_date, data = analysis_data)
-
-# Model 2: pct as a function of end_date and pollster (time trend + pollster effect)
-model_date_pollster <- lm(pct ~ end_date + pollster, data = analysis_data)
-
-# Model 3: pct as a function of end_date, pollster, and is_national (includes national vs. state poll effect)
-model_date_pollster_national <- lm(pct ~ end_date + pollster + is_national, data = analysis_data)
-
-# Model 4: Bayesian model using rstanarm
-model_bayesian <- stan_glm(pct ~ end_date + pollster + is_national,
-                           data = analysis_data,
-                           family = gaussian,
-                           prior = normal(0, 5),
-                           chains = 4, iter = 2000)
-
-#### Compare models ####
-# Calculate AIC and BIC for model comparison
-model_comparison <- tibble(
-  Model = c("model_date", "model_date_pollster", "model_date_pollster_national"),
-  AIC = c(AIC(model_date), AIC(model_date_pollster), AIC(model_date_pollster_national)),
-  BIC = c(BIC(model_date), BIC(model_date_pollster), BIC(model_date_pollster_national))
-)
-
-# Print the model comparison table
-print("Model Comparison Table:")
-print(model_comparison)
-
-
-
-#### Select the best model ####
-# Based on AIC/BIC or cross-validation results, we select the best model
-best_model <- model_date_pollster 
-
-#### Use the best model to predict the winner ####
-# Get predictions for pct using the best model
-analysis_data <- analysis_data %>%
-  mutate(final_prediction = predict(best_model))
-
-# Aggregate predictions by candidate to get the overall predicted support
-final_results <- analysis_data %>%
-  group_by(candidate_name) %>%
[…]

### 2024-11-02 scripts/01-test_simulated_data.R
updated scripts

@@ -1,0 +1,91 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the simulated polling dataset
+# Author: Rohan Alexander
+# Date: 26 September 2024
+# Contact: rohan.alexander@utoronto.ca
+# License: MIT
+# Pre-requisites:
+# - The `tidyverse` package must be installed and loaded
+# - 00-simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `starter_folder` rproj
+
+#### Workspace setup ####
+library(tidyverse)
+
+# Load the simulated dataset
+analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+# Test if the dataset was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+#### Test data ####
+
+# 1. Check if the dataset has 50 rows
+if (nrow(analysis_data) == 50) {
+  message("Test Passed: The dataset has 50 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 50 rows.")
+}
+
+# 2. Check if the dataset has 15 columns
+if (ncol(analysis_data) == 16) {
+  message("Test Passed: The dataset has 16 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 16 columns.")
+}
+
+# 3. Check if 'pollster_id' contains only numeric values
+if (all(is.numeric(analysis_data$pollster_id))) {
+  message("Test Passed: 'pollster_id' contains only numeric values.")
+} else {
+  stop("Test Failed: 'pollster_id' contains non-numeric values.")
+}
+
+# 4. Check if 'sample_size' values are within the expected range (500 to 2000)
+if (all(na.omit(analysis_data$sample_size) >= 500 & na.omit(analysis_data$sample_size) <= 2000)) {
+  message("Test Passed: 'sample_size' values are within the expected range.")
+} else {
+  stop("Test Failed: 'sample_size' values are outside the expected range.")
+}
+
+
+# 6. Check if 'population' contains only valid categories
+valid_populations <- c("Adults", "Likely Voters", "Registered Voters")
+if (all(analysis_data$population %in% valid_populations)) {
+  message("Test Passed: 'population' contains only valid categories.")
+} else {
[…]

### 2024-11-02 scripts/03-clean_data.R

@@ -1,0 +1,206 @@
+#### Preamble ####
+# Purpose: Clean the variables so that the data focuses on donald trump and the variables of interest for further analysis and model building
+# Author: Amy Jin
+# Date: 27 October 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: raw data in the correct directory and corret file name that has the required varibles
+# Assumption: These varibles are workable for data analysis and contributes to model building
+
+#### Workspace setup and loading data  ####
+
+library(tidyverse)
+library(janitor)
+library(arrow) # for parquet
+
+# Load the raw polling data
+raw_data <- read_csv("data/01-raw_data/president_polls.csv") |>
+  clean_names()
+
+
+#### Data Cleaning ####
+raw_data <- raw_data %>%
+  mutate(is_national = if_else(is.na(state), 1, 0)) # Create binary for national polls
+
+
+# Filter data to Harris estimates based on high-quality polls after she declared
+filtered_data <- raw_data |>
+  filter(
+    candidate_name == "Donald Trump",
+    numeric_grade >= 2.5
+    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  mutate(
+    num_trump = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+  )
+
+
+# Select relevant columns for analysis
+filtered_data <- filtered_data %>%
+  select(
+    pollster_id, pollster, state,
+    start_date, end_date, sample_size, population, race_id,
+    cycle, election_date, candidate_name, pct, numeric_grade, methodology,
+    num_trump, is_national
+  )
+na_counts <- filtered_data %>%
+  summarise(across(everything(), ~ sum(is.na(.))))
+filtered_data <- filtered_data %>%
+  mutate(
+    pollster_id = as.numeric(pollster_id), # Ensure pollster_id is numeric
+    sample_size = as.numeric(sample_size), # Ensure sample_size is numeric
+    race_id = as.numeric(race_id), # Ensure race_id is numeric
+    cycle = as.numeric(cycle), # Ensure cycle is numeric
[…]

### 2024-11-02 scripts/04-test_analysis_data.R

@@ -1,0 +1,91 @@
+#### Preamble ####
+# Purpose: Tests simulation and analysis data
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+library(arrow)
+library(here)
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+
+
+#### Preamble ####
+# Purpose: Tests simulation and analysis data
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: tidyverse, testthat, arrow, and here packages installed
+# Any other information needed? Make sure the data file is in the correct directory
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+library(arrow)
+library(here)
+
+# Load the data
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+
+#### Test data ####
+
+# 1. Test if the data was loaded successfully
+test_that("Dataset loads correctly", {
+  expect_true(exists("analysis_data"))
+})
+
+# 2. Test if the data has rows
+test_that("Dataset has rows", {
+  expect_gt(nrow(analysis_data), 0)
+})
+
+# 3. Test if the data has columns
+test_that("Dataset has columns", {
+  expect_gt(ncol(analysis_data), 0)
+})
+
+# 4. Test if there are no missing values in essential columns
+test_that("No missing values in essential columns", {
+  essential_cols <- c("pollster_id", "pollster", "state", "start_date", "end_date", "sample_size")
+  expect_true(all(!is.na(analysis_data[, essential_cols])))
+})
+
+# 5. Test if `sample_size` is always greater than zero
+test_that("Sample size is positive", {
+  expect_true(all(analysis_data$sample_size > 0))
+})
+
+# 6. Test if `start_date` is before `end_date`
+test_that("Start date is before end date", {
+  expect_true(all(as.Date(analysis_data$start_date) < as.Date(analysis_data$end_date)))
+})
+
+# 7. Test if `pct` values are between 0 and 100
+test_that("Percentages are between 0 and 100", {
[…]

### 2024-11-02 scripts/06-model_data.R

@@ -1,0 +1,130 @@
+#### Preamble ####
+# Purpose: Create, compare, and select the best model to predict the 2024 US election result
+# Author: [Your Name]
+# Date: [Update with today's date]
+# Contact: [Your Email]
+# License: MIT
+# Pre-requisites: Ensure all necessary data and libraries are available
+# Any other information needed? [Update with additional context]
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(broom)
+library(ggplot2)
+
+#### Read data ####
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+
+#### Define models ####
+
+# Model 1: pct as a function of end_date (time trend only)
+model_date <- lm(pct ~ end_date, data = analysis_data)
+
+# Model 2: pct as a function of end_date and pollster (time trend + pollster effect)
+model_date_pollster <- lm(pct ~ end_date + pollster, data = analysis_data)
+
+# Model 3: pct as a function of end_date, pollster, and is_national (includes national vs. state poll effect)
+model_date_pollster_national <- lm(pct ~ end_date + pollster + is_national, data = analysis_data)
+
+# Model 4: Bayesian model using rstanarm
+model_bayesian <- stan_glm(pct ~ end_date + pollster + is_national,
+                           data = analysis_data,
+                           family = gaussian,
+                           prior = normal(0, 5),
+                           chains = 4, iter = 2000)
+
+#### Compare models ####
+# Calculate AIC and BIC for model comparison
+model_comparison <- tibble(
+  Model = c("model_date", "model_date_pollster", "model_date_pollster_national"),
+  AIC = c(AIC(model_date), AIC(model_date_pollster), AIC(model_date_pollster_national)),
+  BIC = c(BIC(model_date), BIC(model_date_pollster), BIC(model_date_pollster_national))
+)
+
+# Print the model comparison table
+print("Model Comparison Table:")
+print(model_comparison)
+
+
+
+#### Select the best model ####
+# Based on AIC/BIC or cross-validation results, we select the best model
+best_model <- model_date_pollster 
+
+#### Use the best model to predict the winner ####
+# Get predictions for pct using the best model
+analysis_data <- analysis_data %>%
+  mutate(final_prediction = predict(best_model))
+
+# Aggregate predictions by candidate to get the overall predicted support
+final_results <- analysis_data %>%
+  group_by(candidate_name) %>%
[…]

### 2024-11-02 scripts/clean.R

@@ -1,0 +1,9 @@
+#This script is soley for the purpose of reformatting
+library(styler)
+
+# Reformat an R script
+styler::style_file("scripts/00-simulate_data.R")
+styler::style_file("scripts/01-test_simulated_data.R")
+styler::style_file("scripts/03-clean_data.R")
+styler::style_file("scripts/04-test_analysis_data.R")
+styler::style_file("scripts/06-model_data.R")

### 2024-11-02 scripts/01-test_simulated_data.R
Update 01-test_simulated_data.R

@@ -1,13 +1,14 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the simulated polling dataset
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
 # - The `tidyverse` package must be installed and loaded
 # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Any other information needed? Make sure you are in the `Forcasting-US-Elections` rproj
+
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-02 scripts/04-test_analysis_data.R
Update 04-test_analysis_data.R

@@ -1,11 +1,14 @@
 #### Preamble ####
-# Purpose: Tests simulation and analysis data
+# Purpose: Validate the structure and content of the simulated polling dataset
 # Author: Amy Jin
 # Date: 18 October 2024
-# Contact: amyzh.jin@mail.utoronto,ca
+# Contact: amyzh.jin@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:
+# - The `tidyverse`, `testthat`, `arrow`, and `here` packages must be installed and loaded
+# - The dataset file "analysis_data.parquet" should be located in the "data/02-analysis_data" directory
+# Any other information needed? Make sure you are in the correct project directory before running the script
+
 
 
 #### Workspace setup ####

### 2024-11-02 scripts/06-model_data.R
Update 06-model_data.R

@@ -1,11 +1,15 @@
 #### Preamble ####
-# Purpose: Create, compare, and select the best model to predict the 2024 US election result
-# Author: [Your Name]
+# Purpose: Create, compare, and select the best model to predict the 2024 US election result using different modeling approaches
+# Author: Amy Jin
 # Date: [Update with today's date]
-# Contact: [Your Email]
+# Contact: amyzh.jin@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Ensure all necessary data and libraries are available
-# Any other information needed? [Update with additional context]
+# Pre-requisites:
+# - Ensure the `tidyverse`, `rstanarm`, `broom`, and `ggplot2` packages are installed and loaded
+# - Ensure the data files "analysis_data.parquet" and "harris_data.parquet" are located in the "data/02-analysis_data" directory
+# - Make sure you are in the correct project directory before running the script
+# Any other information needed? The script uses Bayesian modeling and compares models using AIC/BIC metrics
+
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-02 scripts/clean.R
Update clean.R

@@ -1,3 +1,12 @@
+#### Preamble ####
+# Purpose: Reformat R scripts for consistent style and readability
+# Author: Amy Jin
+# Date: [Update with today's date]
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `styler` package must be installed and loaded
+# Any other information needed? Ensure the specified script files are in the correct directory
+
 #This script is soley for the purpose of reformatting
 library(styler)
 

### 2024-11-02 scripts/06-model_data.R
Update 06-model_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Create, compare, and select the best model to predict the 2024 US election result using different modeling approaches
 # Author: Amy Jin
-# Date: [Update with today's date]
+# Date: 2024.11.02
 # Contact: amyzh.jin@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:

### 2024-11-02 paper/paper.qmd
update model section

@@
 
-**Model 4**: A Bayesian model implemented using the `rstanarm` package, incorporating the same predictors as Model 3 but allowing for more robust estimation under uncertainty.
+**Model 4**: A Bayesian model implemented using the `rstanarm` package, incorporating the same predictors as Model 2 but allowing for more robust estimation under uncertainty.
 
@@
 -   **Weaknesses**: Potential oversimplification and assumptions of linearity. Future iterations could explore non-linear effects or hierarchical models.
--   **Final Choice**: We selected **Model 3** (linear model with `end_date`, `pollster`, and `is_national`) for its balance of interpretability and predictive performance.
+-   **Final Choice**: We selected **Model 2** (linear model with `end_date` and `pollster`) for its balance of interpretability and predictive performance.
 The Bayesian model serves as a robustness check, with results presented in the appendix.
@@
 
-**Results**: Model 3 has the lowest AIC and BIC, suggesting it offers the best balance between goodness-of-fit and complexity.
+**Results**: Model 2 has the lowest AIC and BIC, suggesting it offers the best balance between goodness-of-fit and complexity.
 This model includes both time trends and categorical effects for pollsters and national versus state-specific polls.
@@
 
-**Interpretation**: Model 3's residuals suggest that the model adequately captures the data's underlying structure, with no major violations of the linear model assumptions.
+**Interpretation**: Model 2's residuals suggest that the model adequately captures the data's underlying structure, with no major violations of the linear model assumptions.
 

### 2024-11-03 scripts/00-simulate_data.R
simulate data

@@ -1,0 +1,58 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of US elections
+# Author: Amy Jin
+# Date: 18 October 2024
+# Contact: amyzh.jin@mail.utoronto,ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `Forcasting-US-Elections` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(lubridate)
+
+library(tibble)
+
+# Set the seed for reproducibility
+set.seed(106)
+
+# Simulate polling data
+polling_data <- tibble(
+  poll_id = 1:10,
+  pollster = c(
+    "YouGov", "Ipsos", "SurveyUSA", "Morning Consult", "Quinnipiac",
+    "Emerson College", "Marist College", "Rasmussen Reports",
+    "Fox News", "CNN"
+  ),
+  sample_size = sample(800:1600, 10, replace = TRUE),
+  population = sample(c("Likely Voters", "Registered Voters"), 10, replace = TRUE),
+  candidate_name = rep(c("Kamala Harris", "Donald Trump"), 5),
+  pct = sample(40:60, 10, replace = TRUE)
+)
+
+# Simulate sample data
+sample_data <- tibble(
+  pollster_id = sample(1:100, 50, replace = TRUE),
+  pollster = sample(c("Pollster A", "Pollster B", "Pollster C"), 50, replace = TRUE),
+  sponsors = sample(c("Sponsor 1", "Sponsor 2", "Sponsor 3"), 50, replace = TRUE),
+  transparency_score = sample(c(NA, seq(0, 10, by = 0.5)), 50, replace = TRUE),
+  state = sample(c("CA", "TX", "NY", "FL", "PA"), 50, replace = TRUE),
+  start_date = sample(seq.Date(from = as.Date("2024-01-01"), to = as.Date("2024-10-01"), by = "day"), 50, replace = TRUE),
+  end_date = sample(seq.Date(from = as.Date("2024-01-02"), to = as.Date("2024-10-03"), by = "day"), 50, replace = TRUE),
+  sponsor_candidate_party = sample(c("Democratic", "Republican"), 50, replace = TRUE),
+  endorsed_candidate_name = sample(c("Candidate A", "Candidate B"), 50, replace = TRUE),
[…]

### 2024-11-03 scripts/03-clean_data.R
modify variable names

@@ -146,7 +146,7 @@ filtered_data <- raw_data |>
   ) |>
   filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
   mutate(
-    num_trump = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   )
 
 
@@ -156,7 +156,7 @@ filtered_data <- filtered_data %>%
     pollster_id, pollster, state,
     start_date, end_date, sample_size, population, race_id,
     cycle, election_date, candidate_name, pct, numeric_grade, methodology,
-    num_trump, is_national
+    num_harris, is_national
   )
 na_counts <- filtered_data %>%
   summarise(across(everything(), ~ sum(is.na(.))))

### 2024-11-03 paper/paper.qmd
update r package usages

@@
 library(tidyr)
-# Load the arrow package
 library(arrow)
 library(here)
+library(rstanarm)
+library(broom)
+library(loo)
 data = read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
@@
 It provides a comprehensive view of voter preferences through aggregated results from numerous national polls conducted by various polling organizations.
-The dataset was simulated, cleaned, analyzed, and tested using the R programming language [@talia], tidyverse [@citetidyverse], knitr [@citeknitr], ggplot2 [@citeggplot2], gt[@gt].
+The dataset was simulated, cleaned, analyzed, and tested using the R programming language [@talia], tidyverse [@citetidyverse], knitr [@citeknitr], ggplot2 [@citeggplot2] for plots, gt[@gt] for tables, tidyr[@R-tidyr], arrow[@R-arrow] for parquet, here[@R-here], rstanarm[@R-rstanarm], broom[@R-broom], loo[@R-loo], lubridate [@R-lubridate], while tibble [@R-tibble] helped simplify data frame management.
+The testthat package [@R-testthat] was essential for unit testing and ensuring code reliability, and we employed styler [@R-styler] for reformatting and maintaining a consistent code style.
 
@@
 #| echo: false
-library(ggplot2)
 
@@
 
-library(tidyverse)
-library(gt)
-
 # Define the 11 categories
@@
 # Categorizing sample size into small, medium, and large groups
-library(tidyverse)
-library(ggplot2)
 
@@
   )
-library(ggplot2)
 
@@
 
-library(rstanarm)
-
 first_model <-
@@
 ```{r, echo=FALSE, include=FALSE, warning=FALSE}
-library(tidyverse)
-library(broom)
-library(rstanarm)
-library(ggplot2)
-library(loo)
 
@@
   AIC = c(AIC(model_date), AIC(model_date_pollster), AIC(model_date_pollster_national)),
-  BIC = c(BIC(model_date), BIC(model_date_pollster), BIC(model_date_pollster_national))
+  BIC = c(BIC(model_date), BIC(model_date_pollster), BIC(model_date_pollster_national)),
 )
@@
 # Load necessary libraries
-library(gt)
-library(loo)
 

### 2024-11-03 paper/paper.qmd
Updated paper with toc, abstract, draft results and discussion.

@@
 title: "Forecasting the US Election 2024"
-subtitle: "##TODO: result here"
+subtitle: "Modeling the U.S. Polling Data to forecast the outcome of he 2024 U.S. Presidential Elections"
 author: 
@@
 date-format: long
-abstract: "first sentence: specify the general area of the paper and encourage the reader; second sentence: specify the dataset and methods at a general level; third sentence: specify the headline result; and a fourth sentence about implications"
+abstract: "This paper uses polling data to project the outcome of the 2024 U.S.
+Presidential election between Kamala Harris and Donald Trump.
+By employing a multiple linear regression model, we predict each candidate’s level of support across states and estimate the corresponding electoral vote counts.
+Our analysis suggests that Harris has a potential advantage in the Electoral College, highlighting the influence of polling trends and pollster-specific biases on election forecasting.
+These findings underscore the importance of aggregating diverse polling data to create a more balanced and accurate prediction of election results."
 format: pdf
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
-Our results are summarized in @tbl-modelresults.
-
-Results will likely require summary statistics, tables, graphs, images, and possibly statistical analysis or maps. There should also be text associated with all these aspects.
-
-Show the reader the results by plotting them where possible. Talk about them. Explain them. That said, this section should strictly relay results.
-
-Regression tables must not contain stars.
+To project the outcome of the 2024 U.S.
+Presidential election, we utilized a statistical model that leverages polling data from FiveThirtyEight to predict overall support for each candidate.
+Our approach employs a regression model that incorporates poll end dates and pollster-specific adjustments, enabling us to generate a "poll-of-polls" forecast and relay these results as a predictor of election outcomes.
+
+**1.
+Prediction of Polling-Based Support:** Using Model 2, we estimated support levels by regressing polling outcomes on poll end dates and pollster-specific indicators.
+This model was applied to generate predicted support percentages for each candidate over the polling period, accounting for variations across pollsters (see Table 1 for regression output).
+These estimates reflect the predicted level of support for each candidate as the election approaches, adjusted for biases that are characteristic of individual polling organizations.
+
+**2.
+Time-Based Trend Analysis:** By incorporating poll end dates, our model captures temporal trends, allowing us to observe shifts in candidate support over time.
+Figure 1 displays the predicted support levels for each candidate over the final months leading up to the election, highlighting any momentum or shifts that may favor one candidate.
+This visualization provides insight into how public opinion has evolved closer to the election date.
+
+**3.
+Adjustment for Pollster-Specific Biases:** Pollster biases can significantly influence polling averages.
+Our model addresses these variations through the pollster-specific coefficients \( \gamma_j \), which adjust each pollster's results based on historical tendencies.
+Figure 2 illustrates these adjustments, showing how certain pollsters tend to lean in favor of one candidate over another.
+This correction improves the reliability of our aggregated forecast by balancing these systematic biases.
+
+**4.
+Estimation of Overall Support:** Aggregating the predicted support from all polls, we computed an average prediction for each candidate as election day nears.
+This aggregation provides a summary of anticipated national-level support based on polling data and adjusted for timing and pollster effects (see Table 2 for summary statistics of aggregated support levels).
+This result allows for a more precise comparison of each candidate's standing in the final stretch of the campaign.
+
+**5.
+Interpretation of Results:** The analysis suggests that [Candidate A] holds a lead in projected support, with a predicted margin of [X percentage points] over [Candidate B].
+This result implies a favorable position for [Candidate A], assuming that polling trends continue as projected.
+While this forecast is based on aggregate polling data, the adjustments for time trends and pollster effects lend robustness to the model's predictions, offering a refined projection of likely support levels for each candidate.
+
+This methodology effectively combines poll end dates and pollster adjustments, bridging raw polling data with refined predictions of national support.
+By adjusting for bias and temporal effects, our model provides a comprehensive forecast that can inform expectations about the election outcome.
 
@@
 
-## What is done in this paper? {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## What is something that we learn about the world?
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## What is another thing that we learn about the world?
-
-## Weaknesses and next steps
-
-What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?
-
-Weaknesses and next steps should also be included.
+## Summary of Findings
+This study utilizes polling data to project the results of the 2024 U.S.
+Presidential election between Kamala Harris and Donald Trump.
+We developed a multiple linear regression model that predicts state-level polling outcomes by incorporating the end date of each poll and accounting for differences in polling methodology across various organizations.
+By aggregating state-level predictions and mapping them to the U.S.
+Electoral College framework, we were able to estimate the potential electoral vote count for each candidate.
+Our model suggests that Kamala Harris has an advantage in terms of projected electoral votes, signaling a possible pathway to victory based on the current polling trends.
+These findings underscore the effectiveness of using a ‘poll-of-polls’ approach and highlight the importance of controlling for both time trends and pollster-specific biases in forecasting.
+
+## Insights from results
+Several key insights emerge from our analysis.
+First, we observe that poll end dates significantly influence the model’s predictions, indicating that voter preferences are not static but fluctuate over time.
+This temporal sensitivity suggests that the timing of polling data collection is crucial for accuracy, particularly as Election Day approaches.
+Second, the model's adjustment for pollster-specific effects reveals substantial variation among polling organizations, with some systematically showing higher support for one candidate over the other.
+This finding highlights the necessity of accounting for methodological differences between pollsters to avoid skewed predictions.
+By integrating these pollster-specific adjustments, we enhance the model’s robustness and create a more balanced and comprehensive projection of voter support.
+
+## Implications of the results
+The implications of these findings are twofold.
+On one level, our results underscore the volatility of public opinion and the influence of polling timing on the perceived support for each candidate.
+The sensitivity to poll timing means that election forecasts can shift rapidly, particularly in the final weeks before the election.
+This insight could be valuable for political campaigns and analysts, who must consider the timing of major announcements or policy shifts to capture late-breaking trends in voter sentiment.
+On another level, the variation among pollsters suggests that analysts should be cautious when interpreting individual poll results.
+Relying on a single poll or even a small set of polls could lead to biased conclusions; aggregating across multiple polls and accounting for systematic differences is essential for obtaining a balanced view.
+
+## Future Communication and Work to Be Done
+While our model provides a strong foundation for predicting electoral outcomes, there are areas where future research could expand and refine these methods.
+One potential improvement would be to incorporate additional variables, such as economic indicators (e.g., unemployment rates, inflation) or significant political events (e.g., debates, foreign policy crises) that may affect voter sentiment.
+Including real-time data feeds could also allow for dynamic model updates, enabling continuous refinement of predictions as new polls are released and conditions evolve.
+
+Another promising direction for future work involves exploring machine learning techniques, such as ensemble models, that could capture complex interactions between variables and improve predictive power.
+Additionally, understanding the impact of digital influence—particularly social media—on voter behavior might provide further context for shifts in public opinion.
+Expanding the model in this way could offer deeper insights into the drivers of voter sentiment and enhance the accuracy of future election forecasts.
+
+Finally, there is room for improving the communication of model results to the public and stakeholders.
+Visualizations of the model’s uncertainty, for example, could help clarify the inherent unpredictability of election forecasting and avoid overconfident interpretations of the results.
+Regular updates or a publicly accessible dashboard could also foster greater transparency, providing a real-time window into how predictions shift with new data.
+This open communication would not only make election predictions more accessible to the general public but also underscore the value of aggregating diverse data sources in political forecasting.
 

### 2024-11-03 paper/paper.qmd
Added paper.pdf and paper.qmd

@@
 title: "Forecasting the US Election 2024"
-subtitle: "Modeling the U.S. Polling Data to forecast the outcome of he 2024 U.S. Presidential Elections"
+subtitle: "Modeling the U.S. Polling Data to forecast the outcome of the 2024 U.S. Presidential Elections"
 author: 
@@
 
+\newpage
 # Introduction
@@
 
-The remainder of this paper is structured as follows. @sec-data....
+The rest of this paper is organized as follows.
+@sec-data covers the data utilized in this analysis, highlighting key variables and sources, particularly focusing on the quality metrics that influence polling accuracy.
+@sec-model details our modeling approach for each candidate, drawing on insights from recent electoral cycles.
+@sec-results provides our predictions for the Electoral College based on the model outputs.
+@sec-discussion examines the implications of our findings and offers suggestions for future research.
+Finally, the @sec-appendixa and @sec-appendixb assesses YouGov's polling methodology alongside our proposed survey methodology.
 
@@
 We use the polling data to forecast the potential outcomes of the 2024 U.S.
-presidential election between Kamala Harris and Donald Trump.The dataset was obtained from https://projects.fivethirtyeight.com/polls/president-general/2024/national/ or (FiveThirtyEight’s “Poll of Polls” for the 2024 U.S.
-Presidential election [@fivethirtyeight2024]).
+presidential election between Kamala Harris and Donald Trump.The dataset was obtained from FiveThirtyEight’s “Poll of Polls” for the 2024 U.S.
+Presidential election [@fivethirtyeight2024].
 It provides a comprehensive view of voter preferences through aggregated results from numerous national polls conducted by various polling organizations.
+
 The dataset was simulated, cleaned, analyzed, and tested using the R programming language [@talia], tidyverse [@citetidyverse], knitr [@citeknitr], ggplot2 [@citeggplot2] for plots, gt[@gt] for tables, tidyr[@R-tidyr], arrow[@R-arrow] for parquet, here[@R-here], rstanarm[@R-rstanarm], broom[@R-broom], loo[@R-loo], lubridate [@R-lubridate], while tibble [@R-tibble] helped simplify data frame management.
@@
 
-**poll_id:** a unique identifier for each poll conducted pollster: which indicates the organization conducting the poll
-
-**sample_size:** representing the total number of respondents
-
-**population:** specifying the voting group described (e.g., likely voters)
-
-**candidate_name:** the names of the candidates in the poll (e.g., Kamala Harris, Donald Trump)
-
-**pct:** the percentage of the vote or support received by each candidate.
+- **poll_id:** a unique identifier for each poll conducted pollster: which indicates the organization conducting the poll
+- **sample_size:** representing the total number of respondents
+- **population:** specifying the voting group described (e.g., likely voters)
+- **candidate_name:** the names of the candidates in the poll (e.g., Kamala Harris, Donald Trump)
+- **pct:** the percentage of the vote or support received by each candidate.
 
 These variables allow us to explore various dimensions of polling data, such as trends in voter support across different states and the influence of pollster reliability on polling outcomes.
-The table below provides a snapshot of the polling data, displaying the first ten entries.
+@tbl-firstten below provides a snapshot of the polling data, displaying the first ten entries.
 This includes the pollster names, sample sizes, and the percentage of support for each candidate, which can reveal patterns in public opinion and help identify how different organizations may report varying levels of support for Kamala Harris and Donald Trump.
-
-### Response variables:
-
-The main response variable of our analysis is the percentage support (pct) for Donald Trump and Kamala Harris shown in each poll.
-This number tells us what percentage of people said they would vote for each candidate when the survey was taken.
-We use it as the basis of our electoral prediction as it helps us understand how people feel about the candidates.
-
-TODO: descibe
-
-```{r}
-#| label: tbl-pct
-#| tbl-cap: "Histogram of Trump Support Percentage (pct)"
-#| echo: false
-ggplot(data, aes(x = pct)) +
-  geom_histogram(binwidth = 5, fill = "#e31a1c", color = "black") +
-  theme_minimal() +
-  labs(
-    x = "Percentage Support (pct)",
-    y = "Frequency"
-  )
-```
-
-### Predictor variables:
-
-Our model takes into account various variables; their names and descriptions are,
-
-**Pollster**: The name of the polling organization that conducted the poll (e.g., YouGov, RMG Research).
-
-**State:** The U.S. state where the poll was conducted.
-
-**Sample Size:** The total number of respondents that participated in the poll. End Date: The date the poll ended.
-
-These variables were selected and kept due to their evident relationship with the response variable and the significance of it.
-More details of the relationships and model are provided in later sections.
 
@@
 
-### Data Manipulation and cleaning:
-
-The data was cleaned, filtered and mutated to better align with our prediction model.
-The software and packages used have been highlighted in @sec-dataoverview.
-The key steps of our cleaning process were:
-
-Filtered out lower quality polls by only keeping polls that met a set numeric grade threshold (numeric_grade \>= 2.5).
-Filtered out pollsters that had less than 5 entries.
-Filtered out data that was collected before July 21st, 2024.
-Added a variable num_trump to count the number of votes for Donald Trump using pct.
-For data entries where the state was missing, they were categorised as “National”, to standardize the data.
-
-The dataset had several other variables that we did not include in our analysis. These variables were either majorly missing values or not informative enough to be included.
-
-## Summary statistics & Relationships
+### Response variables:
+
+The main response variable of our analysis is the percentage support (pct) for Donald Trump and Kamala Harris shown in each poll.
+This number tells us what percentage of people said they would vote for each candidate when the survey was taken.
+We use it as the basis of our electoral prediction as it helps us understand how people feel about the candidates.
+
+```{r, fig.pos='H'}
+#| label: tbl-pct
+#| tbl-cap: "Histogram of Trump Support Percentage (pct)"
+#| echo: false
+ggplot(data, aes(x = pct)) +
+  geom_histogram(binwidth = 5, fill = "#e31a1c", color = "black") +
+  theme_minimal() +
+  labs(
+    x = "Percentage Support (pct)",
+    y = "Frequency"
+  )
+```
+
+### Predictor variables:
+
+Our model takes into account various variables; their names and descriptions are,
+
+- **Pollster**: The name of the polling organization that conducted the poll (e.g., YouGov, RMG Research).
+- **State:** The U.S. state where the poll was conducted.
+- **Sample Size:** The total number of respondents that participated in the poll. End Date: The date the poll ended.
+
+These variables were selected and kept due to their evident relationship with the response variable and the significance of it.
+More details of the relationships and model are provided in later sections.
+
+## Summary Statistics & Relationships
 
@@
 Mean pct values were calculated for each category, and standard deviation was included to illustrate variability within the poll results.
-The table (@tbl-reliability) and bar chart @fig-reliability showing the average `pct` for each candidate across different levels of pollster reliability.
-Discussion: This analysis highlights whether pollsters with higher reliability scores offer more accurate predictions and if their estimates favor one candidate over the other.
-
-```{r}
+The table (@tbl-reliability) shows the average `pct` for each candidate across different levels of pollster reliability.
+However, the analysis shows the converse of our hypothesis to be true.
+
+```{r, fig.pos='H'}
 #| label: tbl-reliability
@@
 
-```{r}
-#| label: fig-reliability
-#| fig-cap: "Average Support Percentage by Pollster Reliability"
-#| echo: false
-
-# Create a more aesthetic bar chart with alpha and no labels
-ggplot(summary_data, aes(x = reliability, y = mean_pct, fill = candidate_name)) +
-  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black", alpha = 0.8) +
-  geom_errorbar(aes(ymin = mean_pct - sd_pct, ymax = mean_pct + sd_pct), 
-                width = 0.2, 
-                position = position_dodge(0.7)) +
-  labs(
-    subtitle = "Comparison of support percentages for Kamala Harris and Donald Trump",
-    x = "Pollster Reliability",
-    y = "Average Support (%)",
-    fill = "Candidate"
-  ) +
-  scale_fill_manual(values = c("Kamala Harris" = "#1f78b4", "Donald Trump" = "#e31a1c")) + # Custom colors
-  theme_minimal(base_size = 14) +
-  theme(
-    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
-    plot.subtitle = element_text(hjust = 0.5, size = 12),
-    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
-    axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
-    legend.position = "top",
-    legend.title = element_text(face = "bold")
-  )
-
-```
-
 ### Impact of Methodology on Poll Results
@@
 This analysis aimed to determine whether specific methodologies consistently resulted in higher or lower support for either candidate.
-Comparison of candidate vote percentages by polling methodology is demonstrated by [@tbl-methodology] Discussion: We observed potential biases in support based on the polling method, providing insights into which methodologies might offer more reliable forecasts.
-
-```{r}
+Comparison of candidate vote percentages by polling methodology is demonstrated by [@tbl-methodology].
+Evidently Harris has higher support across all platforms, but significantly more for mail-to-web, mail-to-phone and probability panel.
+
+```{r, fig.pos='H'}
 #| label: tbl-methodology
@@
 
-The impact of sample size (sample_size) on the accuracy of polling results was examined.
+The impact of sample size *sample_size* on the accuracy of polling results was examined.
 Larger sample sizes are typically considered more reliable, prompting the categorization of polls into small, medium, and large groups based on sample size.
-For each group, the average percentage of support (pct) for Kamala Harris and Donald Trump was calculated, and variability within each sample size category was analyzed to evaluate the reliability of the results.
-A chart[@fig-samplesize] comparing candidate vote percentages by sample size category.
+For each group, the average percentage of support *pct* for Kamala Harris and Donald Trump was calculated, and variability within each sample size category was analyzed to evaluate the reliability of the results.
+[@fig-samplesize] comparing candidate vote percentages by sample size category.
 Discussion: The results suggest whether larger sample sizes produce more accurate and reliable forecasts, helping us to understand the potential limitations of smaller polls.
 
-```{r}
+```{r, fig.pos='H'}
 #| label: fig-samplesize
@@
 
-### Sponsorship and Bias in Polling
-
-Finally, the potential bias of polls sponsored by partisan organizations was assessed to determine whether such sponsorship influenced support for the associated candidate.
-The vote percentages (pct) for Kamala Harris and Donald Trump were compared in polls sponsored by Democratic-leaning and Republican-leaning organizations (sponsor_candidate_party).
-By calculating the average pct for each candidate within partisan-sponsored polls, this analysis explored whether partisan sponsorship resulted in a systematic overestimation of support for a particular candidate.
-A comparison of the polling results in partisan-sponsored polls, broken down by party affiliation is shown by [@fig-sponsorship].
-Discussion: This analysis reveals potential biases in partisan-sponsored polls and assesses the objectivity of different polling organizations.
-
-```{r}
-#| label: fig-sponsorship
-#| fig-cap: "Vote Percentages by Partisan Sponsorship"
-#| echo: false
-# Calculate the average pct and standard deviation for each candidate by sponsorship type
-sponsorship_summary_data <- data %>%
-  filter(sponsor_candidate_party != "Non-Partisan") %>%  # Focus on partisan-sponsored polls
-  group_by(candidate_name, sponsor_candidate_party) %>%
-  summarise(
-    avg_pct = mean(pct),
-    sd_pct = sd(pct),
-    .groups = 'drop'
-  )
-
-# Create the bar chart comparing vote percentages by sponsorship type
-ggplot(sponsorship_summary_data, aes(x = sponsor_candidate_party, y = avg_pct, fill = candidate_name)) +
-  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black", alpha = 0.8) +
-  geom_errorbar(aes(ymin = avg_pct - sd_pct, ymax = avg_pct + sd_pct), 
-                width = 0.2, 
-                position = position_dodge(0.7)) +
-  labs(
-    x = "Sponsorship (Party Affiliation)",
-    y = "Average Support Percentage (%)",
-    fill = "Candidate"
-  ) +
-  scale_fill_manual(values = c("Kamala Harris" = "#1f78b4", "Donald Trump" = "#e31a1c")) +
-  theme_minimal(base_size = 14) +
-  theme(
-    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
-    plot.subtitle = element_text(hjust = 0.5, size = 12),
-    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
-    axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
-    axis.text.x = element_text(angle = 0, hjust = 0.5),
-    legend.position = "top",
-    legend.title = element_text(face = "bold")
-  )
-
-```
+## Data Manipulation and Cleaning:
+
+The data was cleaned, filtered and mutated to better align with our prediction model.
+The software and packages used have been highlighted in @sec-dataoverview.
+The key steps of our cleaning process were:
+
+- Filtered out lower quality polls by only keeping polls that met a set numeric grade threshold (*numeric_grade* \>= 2.5).
+- Filtered out pollsters that had less than 5 entries.
+- Filtered out data that was collected before July 21st, 2024.
+- Added a variable *num_trump* to count the number of votes for Donald Trump using pct.
+- For data entries where the state was missing, they were categorised as “National”, to standardize the data.
+- Added a binary variable *is_national*, where a value of 1 indicates that the vote was national.
+
+The dataset had several other variables that we did not include in our analysis. These variables were either majorly missing values or not informative enough to be included.
 
@@
 
-# Statistical Model and Analysis
+# Statistical Model and Analysis {#sec-model}
 
@@
 
-#### Model Specifications
+### Model Specifications
 
@@
 
-### We mathematically define these models as follows:
+### Model Definitions
+
+We mathematically define these models as follows:
 
@@
 
-#### Variables and Justification
-
-1.  **Dependent Variable**: ( y_i ) (support percentage, `pct`).
+## Variables and Justification
+
+1. **Dependent Variable**: $y_i$ which refers to support percentage, `pct`.
 2. **Predictors**:
-    -   **(** \text{end\_date} ): Captures time trends, which are essential in understanding shifts in public opinion.
-    -   **(** \text{pollster} ): Accounts for differences in polling methodologies and potential biases.
-    -   **(** \text{is\_national} ): Differentiates between polls conducted at the national level and those specific to individual states, as national polls tend to exhibit different characteristics and sample compositions compared to state polls.
+    -  **end_date:** Captures time trends, which are essential in understanding shifts in public opinion.
+    -  **pollster:** Accounts for differences in polling methodologies and potential biases.
+    -  **is\_national**: Differentiates between polls conducted at the national level and those specific to individual states, as national polls tend to exhibit different characteristics and sample compositions compared to state polls.
 
@@
 
-#### Model Justification
+## Model Justification
 
@@
 
-**Model 1** serves as a baseline, capturing the temporal trend.
-
-**Model 2** addresses variability between pollsters, which is crucial given differences in methodologies and sample recruitment.
-
-**Model 3** adds another layer of nuance by accounting for the national vs. state-specific distinction.
-
-**Model 4** employs Bayesian inference, allowing us to incorporate uncertainty and use prior distributions for more robust predictions.
+- **Model 1** serves as a baseline, capturing the temporal trend.
+- **Model 2** addresses variability between pollsters, which is crucial given differences in methodologies and sample recruitment.
+- **Model 3** adds another layer of nuance by accounting for the national vs. state-specific distinction.
+- **Model 4** employs Bayesian inference, allowing us to incorporate uncertainty and use prior distributions for more robust predictions.
 
@@
 
-#### Assumptions, Limitations, and Circumstances
+## Assumptions, Limitations, and Circumstances
 
@@
 
-#### Software and Model Validation
+## Software and Model Validation
 
@@
 
-**AIC/BIC**: For model selection based on goodness-of-fit and penalization for complexity.
-
-**Cross-validation**: (To be detailed further in the appendix) to evaluate out-of-sample performance.
-
-**Residual Diagnostics**: To check for violations of model assumptions.
-
-**Model Convergence**: Monitored using diagnostics provided by `rstanarm` for the Bayesian model.
+- **AIC/BIC**: For model selection based on goodness-of-fit and penalization for complexity.
+- **Cross-validation**: (To be detailed further in the appendix) to evaluate out-of-sample performance.
+- **Residual Diagnostics**: To check for violations of model assumptions.
+- **Model Convergence**: Monitored using diagnostics provided by `rstanarm` for the Bayesian model.
 
@@
 
-#### Strengths, Weaknesses, and Final Model Choice
+## Strengths, Weaknesses, and Final Model Choice
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-**2.
-Time-Based Trend Analysis:** By incorporating poll end dates, our model captures temporal trends, allowing us to observe shifts in candidate support over time.
-Figure 1 displays the predicted support levels for each candidate over the final months leading up to the election, highlighting any momentum or shifts that may favor one candidate.
-This visualization provides insight into how public opinion has evolved closer to the election date.
+ **2. Time-Based Trend Analysis:** By incorporating poll end dates, our model captures temporal trends, allowing us to observe shifts in candidate support over time.
 
@@
 Adjustment for Pollster-Specific Biases:** Pollster biases can significantly influence polling averages.
-Our model addresses these variations through the pollster-specific coefficients \( \gamma_j \), which adjust each pollster's results based on historical tendencies.
-Figure 2 illustrates these adjustments, showing how certain pollsters tend to lean in favor of one candidate over another.
-This correction improves the reliability of our aggregated forecast by balancing these systematic biases.
+Our model addresses these variations through the pollster-specific coefficients $\gamma_j$, which adjust each pollster's results based on historical tendencies.
 
@@
 Estimation of Overall Support:** Aggregating the predicted support from all polls, we computed an average prediction for each candidate as election day nears.
-This aggregation provides a summary of anticipated national-level support based on polling data and adjusted for timing and pollster effects (see Table 2 for summary statistics of aggregated support levels).
+This aggregation provides a summary of anticipated national-level support based on polling data and adjusted for timing and pollster effects.
 This result allows for a more precise comparison of each candidate's standing in the final stretch of the campaign.
@@
  **5.
-Interpretation of Results:** The analysis suggests that [Candidate A] holds a lead in projected support, with a predicted margin of [X percentage points] over [Candidate B].
-This result implies a favorable position for [Candidate A], assuming that polling trends continue as projected.
+Interpretation of Results:** The analysis suggests that Kamala Harris holds a lead in projected support, with a predicted margin of 2% over Donald Trump.
+This result implies a favorable position for Harris, assuming that polling trends continue as projected.
 While this forecast is based on aggregate polling data, the adjustments for time trends and pollster effects lend robustness to the model's predictions, offering a refined projection of likely support levels for each candidate.
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Appendix 1 {.unnumbered}
+# Appendix A: YouGov's Polling Methodology {#sec-appendix}
 
@@
 
-# Appendix B: Idealized Survey Methodology and Survey Design
+# Appendix B: Idealized Survey Methodology and Survey Design {#sec-appendixb}
 
@@
 
-### Idefitiying Strata:
+### Identifying Strata:
 

### 2024-11-04 paper/paper.qmd
add survey

@@
 
-Include a link to your Google Forms survey here. Survey Link Copy of the Survey Questions
+https://forms.gle/d1pLxx5UNsC9x63s9
 
+```{r,echo=FALSE}
+knitr::include_graphics("survey/1.png")
+knitr::include_graphics("survey/2.png")
+knitr::include_graphics("survey/3.png")
+knitr::include_graphics("survey/4.png")
+knitr::include_graphics("survey/5.png")
+knitr::include_graphics("survey/6.png")
+knitr::include_graphics("survey/7.png")
+knitr::include_graphics("survey/8.png")
+knitr::include_graphics("survey/9.png")
+knitr::include_graphics("survey/10.png")
+
+```
 # Additional data details

### 2024-11-04 paper/paper.qmd
Edit writing for abstract and appendix

@@
 By employing a multiple linear regression model, we predict each candidate’s level of support across states and estimate the corresponding electoral vote counts.
-Our analysis suggests that Harris has a potential advantage in the Electoral College, highlighting the influence of polling trends and pollster-specific biases on election forecasting.
+Our analysis suggests that Harris has a potential advantage in the Electoral College, winning 48.1% support, highlighting the influence of polling trends and pollster-specific biases on election forecasting.
 These findings underscore the importance of aggregating diverse polling data to create a more balanced and accurate prediction of election results."
@@
 Former President Donald Trump represents the Republican party, and President Joe Biden initially led re-election for the Democrats.
-However, Biden has dropped out, and Kamala Harris has taken over as the Democratic nominee [@g24], who is the first Black woman nominee.
-Harris is the Both candidates are focusing their efforts on securing votes by addressing their aspect on social issues.
+However, Biden dropped out, and Kamala Harris took over as the Democratic nominee [@g24], the first Black woman nominee.
+Both candidates are focusing their efforts on securing votes by addressing their aspect on social issues.
 
@@
 For example in 2012, pollsters and data modellers such as Nate Silver have all used survey research and statistical models to successfully predict the result of Barack Obama's victory in the presidential election with consistent and reliable forecasts [@b17].
+In our study, we build on this tradition by leveraging polling data to forecast the outcome of the 2024 U.S.
+Presidential election.
+Using a multiple linear regression model, we predict each candidate’s support across various states.
+This approach not only provides an indication of the potential Electoral College outcome but also illustrates the impact of polling trends and pollster-specific biases on election forecasting.
+By aggregating data from multiple sources, our analysis underscores the value of comprehensive polling and enhancing the accuracy of election predictions.
 
@@
 @sec-data covers the data utilized in this analysis, highlighting key variables and sources, particularly focusing on the quality metrics that influence polling accuracy.
-@sec-model details our modeling approach for each candidate, drawing on insights from recent electoral cycles.
+@sec-model details our modelling approach for each candidate, drawing on insights from recent electoral cycles.
 @sec-results provides our predictions for the Electoral College based on the model outputs.
@@
 We use the polling data to forecast the potential outcomes of the 2024 U.S.
-presidential election between Kamala Harris and Donald Trump.The dataset was obtained from FiveThirtyEight’s “Poll of Polls” for the 2024 U.S.
+presidential election between Kamala Harris and Donald Trump.
+The dataset was obtained from FiveThirtyEight’s “Poll of Polls” for the 2024 U.S.
 Presidential election [@fivethirtyeight2024].
@@
 One of the key reasons for us to select YouGov to dive into it is their innovative use of an advanced statistical model known as Multilevel Regression with Post-stratification (MRP).
-This model allows YouGov to have detailed predictions by using survey data with demographic background information, ensuring the representativeness of broader populations.
+This model allows YouGov to have detailed predictions by using survey data with demographic background information, ensuring the representativeness of broader populations; and accurately estimating voter preferences across diverse groups.
 
@@
 From here, calculate their probability of voting for a specific party if people will vote.
-Combined and reweight responses, to give predictions for candidates [@br24].
+Combined and reweighted responses, to give predictions for candidates [@br24].
 
@@
 
-YouGov's frame of the sample is people in the YouGov panel which use a nonprobability sampling.
+YouGov's frame of the sample is people in the YouGov panel which uses a nonprobability sampling.
+This means not every individual in the population has an equal chance of being chosen; samples are selected based on factors other than randomness [@vehovar2016non].
 Everyone can join the YouGov panel, and YouGov will recruit participants (the sample) to the survey with a rigours process.
@@
 However, this traditional recruitment format will constitute only a small portion of our overall efforts, as advancements in technology have made online recruitment significantly more efficient.
-We plan to use phone outreach as a coplelement when online recruitment does not yield sufficient participants.
+We plan to use phone outreach as a complement when online recruitment does not yield sufficient participants.
 This approach helps avoid inclusion errors by reaching populations without internet access or those who are not engaged in online communities.
